---
title: "R Notebook"
output: html_notebook
---

Using simple functions (no packages)

https://stackoverflow.com/questions/10454973/how-to-create-example-data-set-from-private-data-replacing-variable-names-and-l


```{r}
## A data.frame containing sensitive information
df <- data.frame(
    name = rep(readLines(file.path(R.home("doc"), "AUTHORS"))[9:13], each=2),
    hiscore = runif(10, 99, 100),
    passwd = replicate(10, paste(sample(c(LETTERS, letters), 9), collapse="")))

df
```

```{r}
## A function to anonymise columns in 'colIDs' 
##    colIDs can be either column names or integer indices
anonymiseColumns <- function(df, colIDs) {
    id <- if(is.character(colIDs)) match(colIDs, names(df)) else colIDs
    for(id in colIDs) {
        prefix <- sample(LETTERS, 1)
        suffix <- as.character(as.numeric(as.factor(df[[id]])))
        df[[id]] <- paste(prefix, suffix, sep="")
    }
    names(df)[id] <- paste("V", id, sep="")
    df
}
```


## Anonymize the 1t column

```{r}
## Anonymise it
df2 <- anonymiseColumns(df, c(1))
df2
```

```{r}
## Anonymise it
df3 <- anonymiseColumns(df, c(3))
df3
```





## Another solution
Here is my version of the function. Advantages: no for loops, level labels match variable labels, can be applied to any df, ordered variable names beyond 26 letters, normalization of numeric variables...

Thanks go to:
*  @Tyler Rinker for a solution to using column names in apply functions &
*  @Josh O'Brien for his response to this question

It is available here as a gist.

The data from @Josh O'Brien with a non factor variable

https://stackoverflow.com/a/10458688/5270873


```{r}
   df <- data.frame(
  name = rep(readLines(file.path(R.home("doc"), "AUTHORS"))[9:13], each=2),
  hiscore = runif(10, 99, 100),
  passwd = replicate(10, paste(sample(c(LETTERS, letters), 9), collapse="")))

df$passwd<-as.character(df$passwd)

df
```

```{r}
# the function
anonym<-function(df){
  if(length(df)>26){
    LETTERS<-replicate(floor(length(df)/26),{LETTERS<-c(LETTERS, paste(LETTERS, LETTERS, sep=""))})
    }
    names(df)<-paste(LETTERS[1:length(df)])

    level.id.df<-function(df){
        level.id<-function(i){
      if(class(df[,i])=="factor" | class(df[,i])=="character"){
        column<-paste(names(df)[i],as.numeric(as.factor(df[,i])), sep=".")}else if(is.numeric(df[,i])){
          column<-df[,i]/mean(df[,i], na.rm=T)}else{column<-df[,i]}
          return(column)}
      DF <- data.frame(sapply(seq_along(df), level.id))
      names(DF) <- names(df)
      return(DF)}
    df<-level.id.df(df)
    return(df)}

anonym(df)
```

