---
title: "R Notebook"
output: html_notebook
---


https://www.r-bloggers.com/data-anonymization-in-r/


## Use cases
Public reports.
Public data sharing, e.g. R packages download logs from CRAN's RStudio mirror – cran-logs.rstudio.com – mask ip addresses.
Reports or data sharing for external vendor.
Development works can operate on anonymized PRODUCTION data.
Manually or semi-manually populated data can often brings some new issue after migration to PRODUCTION data.
Such anonymized PRODUCTION data can be quite handy for the devs.

## Dependencies
```{r}
suppressPackageStartupMessages({
  library(data.table)
  library(digest)
  library(knitr) # used only for post creation
})

set.seed(123)
```


## Sample of survey data
Anonymize sensitive information in survey data, data storage in a single table.


```{r}
library(data.table)
# copy data from the web page
# SURV <- read.delim("clipboard")

load(file = "./data/survey.rda")
load(file = "./data/customer.rda")
load(file = "./data/sales.rda")

SURV     <- as.data.table(SURV)
CUSTOMER <- as.data.table(CUSTOMER)
SALES    <- as.data.table(SALES)
```




## Anonymize function
Function will calculate hashes only for unique inputs and return vector of masked inputs.
My version will use digest(x, algo="crc32") because it fits better into html tables, algo crc32 is not really secure.

Read ?digest::digest for supported algo, also consider to salt your input vector, e.g. x=paste0("prefix",x,"suffix"). Performance improvement possible using Rcpp / C: digest #2.

```{r}
# serializing requires that the variables or columns are characters not factors
anonymize <- function(x, algo = "crc32") {
    unq_hashes <- vapply(unique(x), function(object)
                digest(object, algo = algo, serialize = FALSE), FUN.VALUE = "", USE.NAMES = TRUE)
    unname(unq_hashes[x])
}
```


## Anonymize survey data
We will keep city and sex fields unmasked.

```{r}
# choose columns to mask
cols_to_mask <- c("NAME", "ADDRESS", "POSTAL.CODE")
# backup original data
SURV_ORIG <- copy(SURV)
# anonymize
SURV[, (cols_to_mask) := lapply(.SD, anonymize),.SDcols=cols_to_mask]
# pretty print
head(SURV)
```

## Why not just random data or integer sequence
When using the digest function to hide sensitive data you:
* keep rows distribution: aggregates by masked columns will still match to aggregates on original columns, see simple grouping below:

```{r}
SURV_ORIG[, .(.N, 
             mean_age = mean(AGE), 
             mean_score = mean(SCORE)), 
         by = .(CITY, POSTAL.CODE)][, .SD]
```

We do the same for the modified data:

```{r}
SURV[, .(.N, 
             mean_age = mean(AGE), 
             mean_score = mean(SCORE)), 
         by = .(CITY, POSTAL.CODE)][, .SD]
```

* keep relationships on equi joins: `if t1.col1 == t2.col4 TRUE` then also `digest(t1.col1) == digest(t2.col4) TRUE`.
Example in next section below.


## Sample of sales data
Anonymize relational data in sales data, data normalized into SALES and CUSTOMER tables.

```{r}
head(SALES, 4)
```

```{r}
head(CUSTOMER, 2)
```



```{r}
# CUSTOMER[SALES, on="CUSTOMER.UID"]
setkey(CUSTOMER, CUSTOMER.UID)                 # set the key on CUSTOMER.UID
setkey(SALES, CUSTOMER.UID) 
CUSTOMER[SALES]
```



```{r}
# join and aggregate. Group by city and postal code
CUSTOMER[SALES][, .(QUANTITY = sum(QUANTITY), 
                    VALUE = sum(VALUE)), 
                by = .(CITY, POSTAL.CODE)]
```


## Anonymize sales data

```{r}
SALES[, CUSTOMER.UID := anonymize(CUSTOMER.UID)]

cols_to_mask <- c("CUSTOMER.UID", "NAME", "ADDRESS", "POSTAL.CODE")
# cols_to_mask <- c("NAME", "ADDRESS", "POSTAL.CODE")
CUSTOMER[, (cols_to_mask) := lapply(.SD, anonymize), .SDcols = cols_to_mask]
setkey(CUSTOMER, CUSTOMER.UID)
```



```{r}
CUSTOMER
```


```{r}
SALES
```

```{r}
CUSTOMER[SALES]
```

```{r}
# also the aggregates on masked columns will match to the origin
head(
    CUSTOMER[SALES][,.(QUANTITY = sum(QUANTITY), VALUE = sum(VALUE)), by=.(CITY, POSTAL.CODE)]
)
```
















```{r}
# optional if file has not been saved
# save(SURV, file = "./data/survey.rda")
```

```{r}
# CUSTOMER <- read.csv("customer.csv")
# save(CUSTOMER, file = "./data/customer.rda")
```


```{r}
# save(SALES, file = "./data/sales.rda")
# save(CUSTOMER, file = "./data/customer.rda")
```



```{r}
# optional if survey.rda does not exist
#   x <- SURV <- read.delim("clipboard")
```

```{r}
FactorsToCharacter <- function(bob) {
    newbob_char <- as.data.frame(lapply(bob[sapply(bob, is.factor)], as.character),
                             stringsAsFactors = FALSE)
    newbob_rest <- bob[!(sapply(bob, is.factor))]
    newbob <- cbind(newbob_char, newbob_rest)
    return(newbob)
}
```


```{r}
# SALES <- FactorsToCharacter(SALES)
# str(SALES)
```


```{r}
# CUSTOMER <- FactorsToCharacter(CUSTOMER)
# str(CUSTOMER)
```