---
title: "Biobase 02: S4 classes - read a eSet data file"
output: html_notebook
---

Source: https://www.ebi.ac.uk/arrayexpress/help/expressionset_objects.html
Doc: http://www.bioconductor.org/packages/release/bioc/vignettes/ArrayExpress/inst/doc/ArrayExpress.pdf


ExpressionSet is the de facto object defined in the Bioconductor Biobase package for loading and manipulating microarray data in R. These objects are created using the ArrayExpress R package using default parameters. For each experiment, raw (un-normalized) data files are stored in the object as well as other metadata from its MAGE-TAB files, such as study, sample and gene metadata. Please refer to the package ArrayExpress package vignette for more details.

To load the E-MTAB-777.eSet.r object, download it, start R, and run the following command:

```{r}
load("D:/Dropbox/git/extdata/E-MTAB-777.eSet.r")
```

This will create an ExpressionSet object named "study". If you plan to load multiple ExpressionSet objects, you should rename this object before loading the next one to prevent it from being overwritten. For example:

```{r}
mtab777eSet <- study
```


```{r}
class(mtab777eSet)
```

```{r}
print(mtab777eSet)
```

```{r}
phenoData(mtab777eSet)
```

```{r}
featureData(mtab777eSet)
```

```{r}
experimentData(mtab777eSet)
```


```{r}
annotation(mtab777eSet)
```

```{r}
mtab777eSet$gender[1:5]
```

```{r}
featureNames(mtab777eSet)[1:5]
```


```{r}
varLabels(mtab777eSet)[1:5]
```

```{r}
 mat <- exprs(mtab777eSet)
 dim(mat)
```

```{r}
mtab777eSet
```

