---
title: "R Notebook"
output: html_notebook
---

https://gis.stackexchange.com/questions/18940/converting-point-to-lat-lon


## Projecting sp objects in R 2 answers
I have received a shapefile with measurement points to be analyzed, but no matter how I try, I can't find how to convert to lat lon. I suppose it would be good to know the projection but when I look into the file (with R and summary()) it says "Is projected: NA" and "proj4string : [NA]".

One measurement-point I know is close to the border Peru-Chile has x,y = 354521, 7997417.8.

How can I find out how to convert this point (and many others) to lat lon?

I tried to follow the instructions of similar questions here, but to no avail.



```{r}
library(proj4)
proj4string <- "+proj=utm +zone=19 +south +ellps=WGS84 +datum=WGS84 +units=m +no_defs "

# Source data
xy <- data.frame(x=354521, y=7997417.8)

# Transformed data
pj <- project(xy, proj4string, inverse=TRUE)
latlon <- data.frame(lat=pj$y, lon=pj$x)
print(latlon)

# -18.10714	-70.37495	
```

```{r}
# Malaysian coordinates

library(proj4)
proj4string <- "+proj=utm +zone=47 +north +ellps=WGS84 +datum=WGS84 +units=m +no_defs "

# Source data
xy <- data.frame(x = 801553.296, y = 349423.677)

# Transformed data
pj <- project(xy, proj4string, inverse=TRUE)
latlon <- data.frame(lat=pj$y, lon=pj$x)
print(latlon)

# -18.10714	-70.37495	
```

```{r}
latlon$lat
```

