---
title: "HDF5 rhdf5 package discovery"
output: html_notebook
---


## 1st HDF5 example: Creating file and groups

```{r}
# Create an empty HDF5 file
library(rhdf5)
h5createFile("myhdf5file.h5")
```


The HDF5 file can contain a group hierarchy. We create a number of groups and list the file content afterwards.

```{r}
# create group foo
h5createGroup("myhdf5file.h5","foo")
```


```{r}
# create group baa
 h5createGroup("myhdf5file.h5","baa")
```


```{r}
# create sub group under foo
h5createGroup("myhdf5file.h5","foo/foobaa")
```


```{r}
# list what's in it
h5ls("myhdf5file.h5")
```



### 2nd example: writing and reading objects

```{r}
# writing a matrix
A = matrix(1:10, nr=5, nc=2)
h5write(A, "myhdf5file.h5","foo/A")

# writing an array
B = array(seq(0.1,2.0,by=0.1), dim = c(5, 2, 2))    # 5x2x2 array
attr(B, "scale") <- "liter"
h5write(B, "myhdf5file.h5","foo/B")

# writing an array of characters
C = matrix(paste(LETTERS[1:10], LETTERS[11:20],     # matrix 2x5
                 collapse=""), nr=2,nc=5)
h5write(C, "myhdf5file.h5","foo/foobaa/C")

# writing a dataframe
df = data.frame(1L:5L, 
                seq(0,1, length.out=5), 
                c("ab","cde","fghi","a","s"), stringsAsFactors=FALSE)

h5write(df, "myhdf5file.h5", "df")     

h5ls("myhdf5file.h5")
```

> with write we have to specify the object and the name of the object.


### write a dataframe with column names
```{r}
# writing a dataframe
df2 = data.frame(long.int = 1L:5L, 
                sequence =seq(0,1, length.out=5), 
                strings = c("ab","cde","fghi","a","s"), 
                stringsAsFactors=FALSE)

h5write(df2, "myhdf5file.h5","df2")
h5ls("myhdf5file.h5")
```

### saving individual objects wit `h5save`
```{r}
well.name <- "EXAMPLE_01"
well.wh <- 100
well.bh <- 6000

h5save(well.name, well.bh, well.wh, file = "myhdf5file.h5")
h5ls("myhdf5file.h5")
```

> with `h5save` we do not need to specify the name of the object


### saving a list that combines different data types

```{r}
H5Fopen("myhdf5file.h5")

# a list is saved like a group

well.input <- list(oil.api = 40, 
                oil.visc = 5,
                gas.sg = 0.65,
                liq.rt = 1000,
                well.name = "well_01",
                segments = as.integer(30),
                deltaL = 1:40
                )

h5save(well.input, file = "myhdf5file.h5")
h5ls("myhdf5file.h5")
H5close()
```

### saving with `h5write`
```{r}
H5Fopen("myhdf5file.h5")

# a list is saved like a group
# each of the objects is a scalar
# the list or group has to be removed first in order to write it
# the list or group does not allow overwritting
# the list or group supports sequences. they are written as 1-col vectors


well.input2 <- list(oil.api = 40, 
                oil.visc = 5,
                gas.sg = 0.65,
                liq.rt = 1000,
                well.name = "well_01",
                segments = as.integer(50),
                deltaL = 1:50,
                water.sg = 1.01
                )

h5write(well.input2, file = "myhdf5file.h5", name = "well.input2")
h5ls("myhdf5file.h5")
H5close()
```

### saving a list of scalars and a dataframe
```{r}
H5Fopen("myhdf5file.h5")

# writing a dataframe
df3 = data.frame(long.int = 1L:5L, 
                sequence = seq(0,1, length.out=5), 
                strings = c("ab","cde","fghi","a","s"), 
                stringsAsFactors=FALSE)

well.input3 <- list(oil.api = 40, 
                oil.visc = 5,
                gas.sg = 0.65,
                liq.rt = 1000,
                well.name = "well_01",
                segments = as.integer(50),
                deltaL = 1:50,
                water.sg = 1.01,
                dataframe = df3
                )

h5write(well.input3, file = "myhdf5file.h5", name = "well.input3")
h5ls("myhdf5file.h5")
H5close()
```


### writting attributes to group or list with `h5write`
```{r}
H5Fopen("myhdf5file.h5")

# writing a dataframe
df3 = data.frame(long.int = 1L:5L, 
                sequence = seq(0,1, length.out=5), 
                strings = c("ab","cde","fghi","a","s"), 
                stringsAsFactors=FALSE)

well.input4 <- list(oil.api = 40, 
                oil.visc = 5,
                gas.sg = 0.65,
                liq.rt = 1000,
                well.name = "well_01",
                segments = as.integer(50),
                deltaL = 1:50,
                water.sg = 1.01,
                dataframe = df3
                )

h5write(well.input4, 
        file = "myhdf5file.h5", 
        name = "well.input4",
        write.attributes = TRUE
        )
h5ls("myhdf5file.h5")
H5close()
```

> When write.attributes is TRUE, the group shows a red "A" letter.
The default attributes of the list are shown as name = "names" and values are the names of the members of the list.



```{r}
h5ls("myhdf5file.h5")
```


```{r}
fid <- H5Fopen("myhdf5file.h5")
# writing a matrix
A = matrix(1:10, nr=5, nc=2)

# writing an array
B = array(seq(0.1,2.0,by=0.1), dim = c(5, 2, 2))    # 5x2x2 array

# writing a dataframe
df = data.frame(1L:5L, 
                seq(0,1, length.out=5), 
                c("ab","cde","fghi","a","s"), stringsAsFactors=FALSE)

#h5createDataset("myhdf5file.h5", "COMP", c(5,8))

h5writeDataset.data.frame(df, h5loc = fid, name = "COMP1", DataFrameAsCompound = TRUE)


```



```{r}
# create new dataset
h5createDataset("myhdf5file.h5", "foo/S", c(5,8), 
                storage.mode = "integer", chunk=c(5,1), level=7)
h5ls("myhdf5file.h5")
```

```{r}
h5write(matrix(1:5, nr=5, nc=1), file="myhdf5file.h5",         # matrix 5x1
         name="foo/S", index=list(NULL,1))                     # it will write only a vector of S
h5ls("myhdf5file.h5")
```

```{r}
# we can see here that only the first column of the matrix S was written
h5read("myhdf5file.h5", "foo/S")
```

```{r}
h5write(6:10, file="myhdf5file.h5",              # write the 1st row of S, columns 2 thru 6
        name="foo/S", index = list(1,2:6))

h5read("myhdf5file.h5", "foo/S")
```

```{r}
# write data in S on rows 1 thru 5 and columns 3 thru 8
h5write(matrix(11:40, nr=5, nc=6), file="myhdf5file.h5",
        name="foo/S", index=list(1:5, 3:8))

h5read("myhdf5file.h5", "foo/S")
```

```{r}
# write a sequence in rows 3,4 and columns 1,2
h5write(matrix(141:144, nr=2, nc=2), file="myhdf5file.h5", 
        name="foo/S", index=list(3:4, 1:2))

h5read("myhdf5file.h5", "foo/S")
```

```{r}
# write sequence 2x2 in rows 2, 3and columns 3 and 6 only
h5write(matrix(151:154,nr=2,nc=2), file="myhdf5file.h5", 
        name="foo/S", index=list(2:3, c(3,6)))
h5read("myhdf5file.h5", "foo/S")
```

```{r}
# read rows 2:3 and columns 2:3
h5read("myhdf5file.h5", "foo/S", index=list(2:3, 2:3))
```

```{r}
# read specific columns
h5read("myhdf5file.h5", "foo/S", index=list(2:3,c(2,4)))
```

```{r}
# read 4 specified columns
h5read("myhdf5file.h5", "foo/S", index=list(2:3,c(1,2,4,5)))
```

###

```{r}
# create another dataset H with 5x8 
h5createDataset("myhdf5file.h5", "foo/H", c(5,8), 
                storage.mode = "integer",
                chunk=c(5,1), level=7)
h5read("myhdf5file.h5", "foo/H")
```

```{r}
h5write(matrix(1:5,nr=5,nc=1), file="myhdf5file.h5", name="foo/H", 
        start=c(1,1))
h5read("myhdf5file.h5", "foo/H")
```


```{r}
h5write(6:10, file="myhdf5file.h5", name="foo/H",
        start=c(1,2), count=c(1,5))
h5read("myhdf5file.h5", "foo/H")
```

```{r}
h5write(matrix(11:40,nr=5,nc=6), file="myhdf5file.h5", name="foo/H",
        start=c(1,3))
h5read("myhdf5file.h5", "foo/H")
```


```{r}
h5write(matrix(141:144,nr=2,nc=2), file="myhdf5file.h5", name="foo/H", 
          start=c(3,1))
 h5read("myhdf5file.h5", "foo/H")
```

```{r}
# write small matrix in two different columns
h5write(matrix(151:154,nr=2,nc=2), file="myhdf5file.h5", name="foo/H",
        start=c(2,3), stride=c(1,3))

h5read("myhdf5file.h5", "foo/H")
```

```{r}
h5read("myhdf5file.h5", "foo/H", 
       start=c(2,1),             # start at row=2, col=1
       stride=c(1,3),            # only columns 1 and 3
       count=c(2,2),             # 
       block=c(1,2))
```


```{r}
h5ls("myhdf5file.h5", all=TRUE)
```


```{r}
h5ls("myhdf5file.h5", recursive=2)
```


## the attribute interface
```{r}
H5close()

aFile <- file.path("./", "ex_H5A.h5")
if (file.exists(aFile)) {
    file.remove(aFile)
}

# create a file and write something
h5createFile(aFile)
h5write(1:15, aFile, "A")

# write an attribute 'unit' to 'A'
fid <- H5Fopen(aFile)        # open file
did <- H5Dopen(fid, "A")           # open dataset
sid <- H5Screate_simple(c(1,1))    # create the dataspace

# datatype setting
tid <- H5Tcopy("H5T_C_S1")         # copy the datatype string
H5Tset_size(tid, 10L)              # string of size 10

# create the attribute passing 
#                dataset, name of attribute, dataspace, datatype
aid <- H5Acreate(did,     "unit",            tid,       sid)
cat("\nThis is the attribute id \n")
aid

# write the value of the attribute
H5Awrite(aid, "liter")

cat("\nThis is the attribute id and value \n")
res <-  H5Aread(aid)
res

H5Aclose(aid)         # close
H5Sclose(sid)         # close

# check if attribute exist
cat("\ncheck if attribute exist \n")
H5Aexists(did, "unit")
# [1] TRUE


H5Dclose(did)
H5Fclose(fid)
cat("\n dump the file\n")
h5dump(aFile)
```

## the dataset interface

```{r}

aFile <- file.path("./", "ex_H5D.h5")
if (file.exists(aFile)) {
    file.remove(aFile)
}

# write a dataset
fid <- H5Fcreate(aFile)
fid
sid <- H5Screate_simple(c(10,5,3))
sid
did <- H5Dcreate(fid, "A", "H5T_STD_I32LE", sid)
did
H5Dwrite(did, 1L:150L, h5spaceMem = sid, h5spaceFile = sid)
H5Dclose(did)
H5Sclose(sid)
H5Fclose(fid)

# read a dataset
fid <- H5Fopen(aFile)
fid
did <- H5Dopen(fid, "A")
did
sid <- H5Dget_space(did)
sid
B <- H5Dread(did)
B
H5Dclose(did)
H5Sclose(sid)
H5Fclose(fid)

# write a subarray
fid <- H5Fopen(aFile)
fid
did <- H5Dopen(fid, "A")
did
sid <- H5Dget_space(did)
sid
H5Sselect_index(sid, list(1:3,2:4,2))
sidmem <- H5Screate_simple(c(3,3,1))
sidmem
A = array(-801:-809,dim=c(3,3,1))
H5Dwrite(did, A, h5spaceMem = sidmem, h5spaceFile = sid)
H5Dread(did)
H5Sclose(sid)
H5Dclose(did)
H5Sclose(sidmem)
H5Fclose(fid)
```

## the file interface
```{r}
fid <- H5Fcreate("ex_H5F.h5")
fid
H5Fclose(fid)

fid2 <- H5Fopen("ex_H5F.h5")
fid2
H5Fclose(fid2)
```

## the group interface

```{r}
aFile <- file.path("./", "ex_H5G.h5")
if (file.exists(aFile)) {
    file.remove(aFile)
}

fid <- H5Fcreate(aFile)
gid <- H5Gcreate(fid, "foo")
gid

H5Gget_info(gid)
H5Gclose(gid)

H5Gget_info_by_idx(fid,1)
H5Gget_info_by_name(fid,"foo")

H5Fclose(fid)
```

## the link interface

```{r}
aFile <- file.path("./", "ex_H5L.h5")
if (file.exists(aFile)) file.remove(aFile)

# create an hdf5 file and a group
h5createFile(aFile)
h5createGroup(aFile, "foo")

# reopen file and get link info
fid <- H5Fopen(aFile)
cat("\n Does foo link exist? \n")
H5Lexists(fid, "foo")
cat("\n Does baa link exist? \n")
H5Lexists(fid, "baa")

cat("\nGet info \n")
H5Lget_info(fid, "foo")
H5Fclose(fid) 
```

## the object interface
```{r}
aFile <- file.path("./", "ex_H5O.h5")
if (file.exists(aFile)) file.remove(aFile)

# create an hdf5 file and write something
h5createFile(aFile)
h5createGroup(aFile,"foo")
B = array(seq(0.1,2.0,by=0.1),dim=c(5,2,2))
h5write(B, aFile, "foo/B")

# reopen file and dataset and get object info
fid <- H5Fopen(aFile)
oid <- H5Oopen(fid, "foo")

H5Oget_num_attrs(oid)
H5Oclose(oid)
H5Fclose(fid)
```

## Create attributes

```{r}
H5close()
aFile <- file.path("./", "ex_createAttribute.h5")
if (file.exists(aFile)) file.remove(aFile)

h5createFile(aFile)
h5write(1:1, aFile, "A")
fid <- H5Fopen(aFile)
did <- H5Dopen(fid, "A")
h5createAttribute (did, "time", c(1,10))
h5createAttribute (did, "units", 0)
h5createAttribute (did, "width", 100)
did
h5ls(aFile)
H5Dclose(did)
H5Fclose(fid)
```

