---
title: "R Notebook"
output: html_notebook
---


```{r}
#Install rhdf5 library
#source("http://bioconductor.org/biocLite.R")
#biocLite("rhdf5")


library("rhdf5")
#also load ggplot2 and dplyr
library("ggplot2")
library("dplyr")
library("scales")
```

### HDF5 Quick Review

The HDF5 format is a self-contained directory structure. In HDF5 files though “directories” are called “groups” and “files” are called “datasets”. Each element in an hdf5 file can have metadata attached to it making HDF5 files “self-describing”.

Read more about HDF5 here.

### Explore the HDF5 File Structure

Let’s first explore an HDF5 file that we know nothing about using the R function, h5ls.

```{r}
#NOTE: be sure to adjust the path to match your file structure!
# Identify file path
f <- "NEONDSTowerTemperatureData.hdf5"
# View structure of file
h5ls(f)
```

Note that h5ls returns the structure of the HDF5 file structure including the group and dataset names and associated types and sizes of each object. In our file, there are datasets that are compound in this file. Compound class means there are a combination of datatypes within the datasets (e.g. numbers, strings, etc)
contained within that group.

Also note that you can add the recursive variable to the h5ls command to set the number of nested levels that the command returns. Give it a try.

```{r}
#specify how many "levels" of nesting are returns in the command
h5ls(f, recursive=2)

##        group      name     otype dclass dim
## 0          / Domain_03 H5I_GROUP           
## 1 /Domain_03      OSBS H5I_GROUP           
## 2          / Domain_10 H5I_GROUP           
## 3 /Domain_10      STER H5I_GROUP
```


```{r}
h5ls(f, recursive=3)

##             group      name     otype dclass dim
## 0               / Domain_03 H5I_GROUP           
## 1      /Domain_03      OSBS H5I_GROUP           
## 2 /Domain_03/OSBS     min_1 H5I_GROUP           
## 3 /Domain_03/OSBS    min_30 H5I_GROUP           
## 4               / Domain_10 H5I_GROUP           
## 5      /Domain_10      STER H5I_GROUP           
## 6 /Domain_10/STER     min_1 H5I_GROUP           
## 7 /Domain_10/STER    min_30 H5I_GROUP
```

## ## About the structure

Looking at the h5ls output, we see this H5 file has a nested group and dataset structure. Below, we will slice out temperature data which is located within the following path: Domain_03 --> OSBS --> min_1 --> boom_1 --> temperature

Take note that this path is 4 groups “deep” and leads to one dataset called temperature in this part of the HDF5 file as follows:

* Domain_03 - A NEON domain is an ecologically unique region. Domain 3 is one of 20 regions that NEON uses to organize its network spatially .

* OSBS - a group representing data from the Ordway Swisher Biological Station.

* min_1 - A group representing the mean temperature data value for every for one minute in time. Temperature data is often collected at high frequencies (20 hz or 20 measurements a second) or more. A typical data product derived from high frequency tempearture data is an average value. In this case, all measurements are averaged every minute.

* boom_1 - Boom 1 is the first and lowest arm or level on the tower. Towers often contain arms where the sensors are mounted, that reach out horizontally away from the tower (see figure below). The tower at Ordway Swisher has a total of 6 booms (booms 1-5 and the tower top).

A NEON tower contains booms or arms that house sensors at varying heights along the tower.

```{r}
#read in temperature data
temp <- h5read(f, "/Domain_03/OSBS/min_1/boom_1/temperature")
#view the first few lines of the data 
head(temp)

##                    date numPts     mean      min      max    variance
## 1 2014-04-01 00:00:00.0     60 15.06154 14.96886 15.15625 0.002655015
## 2 2014-04-01 00:01:00.0     60 14.99858 14.93720 15.04274 0.001254117
## 3 2014-04-01 00:02:00.0     60 15.26231 15.03502 15.56683 0.041437537
## 4 2014-04-01 00:03:00.0     60 15.45351 15.38553 15.53449 0.001174759
## 5 2014-04-01 00:04:00.0     60 15.35306 15.23799 15.42346 0.003526443
## 6 2014-04-01 00:05:00.0     60 15.12807 15.05846 15.23494 0.003764170
##        stdErr uncertainty
## 1 0.006652087  0.01620325
## 2 0.004571866  0.01306111
## 3 0.026279757  0.05349682
## 4 0.004424851  0.01286833
## 5 0.007666423  0.01788372
## 6 0.007920616  0.01831239
```

```{r}
#generate a quick plot of the data, type=l for "line"
plot(temp$mean,type='l')
```

We can make our plot look nicer by adding date values to the x axis. However, in order to list dates on the X axis, we need to assign the date field a date format so that R knows how to read and organize the labels on the axis.

```{r}
# Let's clean up the plot abovet. We can first add dates to the x axis. 
# In order to list dates, we need to specify the format that the date field is in.
# First read in the time as UTC format
temp$date <- as.POSIXct(temp$date ,format = "%Y-%m-%d %H:%M:%S", tz = "GMT")

# Create a plot using GGPLOT
ordwayPlot <- ggplot(temp,aes(x=date,y=mean))+
  geom_path()+
  ylab("Mean temperature") + 
  xlab("Date in UTC")+
  ggtitle("3 Days of Temperature Data at Ordway Swisher")

ordwayPlot
```

Let’s have a close look at this plot. Notice anything unusual with it? HINT: when would you expect the temperature to be the warmest during the day?

In many parts of the world, we'd expect the maximum temperature to occur during late afternoon.
In this case, our data are in UTC time zone. The UTC time zone is a standardized time zone that does not observe daylight savings time. If your data are in UTC, then you will need to CONVERT them to your local time zone for the dates and times to make sense when you plot / analyze your data.

For example, find the time zone for Denver, Colorado in this wikipedia page. How many hours difference is UTC from the local time in Denver, Colorado?

To adjust for time, we need to tell R that the time zone is eastern standard. We can use the attributes function to set the time zone.


```{r}
#convert to local time for pretty plotting
attributes(temp$date)$tzone <- "America/New_York"

ordwayPlot2 <- ggplot(temp,aes(x=date,y=mean))+
  geom_path()+
  ylab("Mean temperature") + xlab("Date in Eastern Standard Time")+
  theme_bw()+
  ggtitle("3 Days of Temperature Data at Ordway Swisher")+
  scale_x_datetime(breaks=pretty_breaks(n=4))

#let's check out the plot
ordwayPlot2
```

## Extracting metadata
Metadata can be stored directly within HDF5 files and attached to each group or 
dataset in the file - or to the file itself. To read the metadata for elements
in a HDF5 file in R we use the h5readAttributes function.

```{r rows.print=40}
## View the groups and datasets in our file, 
#we will grab the nested structure, 5 'levels' down
#5 levels gets us to the temperature dataset
fiu_struct <- h5ls(f, recursive = 5)
class(fiu_struct)
## have a look at the structure.
fiu_struct

##                               group        name       otype   dclass  dim
## 0                                 /   Domain_03   H5I_GROUP              
## 1                        /Domain_03        OSBS   H5I_GROUP              
## 2                   /Domain_03/OSBS       min_1   H5I_GROUP              
## 3             /Domain_03/OSBS/min_1      boom_1   H5I_GROUP              
## 4      /Domain_03/OSBS/min_1/boom_1 temperature H5I_DATASET COMPOUND 4323
## 5             /Domain_03/OSBS/min_1      boom_2   H5I_GROUP              
## 6      /Domain_03/OSBS/min_1/boom_2 temperature H5I_DATASET COMPOUND 4323
## 7             /Domain_03/OSBS/min_1      boom_3   H5I_GROUP              
## 8      /Domain_03/OSBS/min_1/boom_3 temperature H5I_DATASET COMPOUND 4323
## 9             /Domain_03/OSBS/min_1      boom_5   H5I_GROUP              
## 10     /Domain_03/OSBS/min_1/boom_5 temperature H5I_DATASET COMPOUND 4323
## 11            /Domain_03/OSBS/min_1   tower_top   H5I_GROUP              
## 12  /Domain_03/OSBS/min_1/tower_top temperature H5I_DATASET COMPOUND 4323
## 13                  /Domain_03/OSBS      min_30   H5I_GROUP              
## 14           /Domain_03/OSBS/min_30      boom_1   H5I_GROUP              
## 15    /Domain_03/OSBS/min_30/boom_1 temperature H5I_DATASET COMPOUND  147
## 16           /Domain_03/OSBS/min_30      boom_2   H5I_GROUP              
## 17    /Domain_03/OSBS/min_30/boom_2 temperature H5I_DATASET COMPOUND  147
## 18           /Domain_03/OSBS/min_30      boom_3   H5I_GROUP              
## 19    /Domain_03/OSBS/min_30/boom_3 temperature H5I_DATASET COMPOUND  147
## 20           /Domain_03/OSBS/min_30      boom_5   H5I_GROUP              
## 21    /Domain_03/OSBS/min_30/boom_5 temperature H5I_DATASET COMPOUND  147
## 22           /Domain_03/OSBS/min_30   tower_top   H5I_GROUP              
## 23 /Domain_03/OSBS/min_30/tower_top temperature H5I_DATASET COMPOUND  147
## 24                                /   Domain_10   H5I_GROUP              
## 25                       /Domain_10        STER   H5I_GROUP              
## 26                  /Domain_10/STER       min_1   H5I_GROUP              
## 27            /Domain_10/STER/min_1      boom_1   H5I_GROUP              
## 28     /Domain_10/STER/min_1/boom_1 temperature H5I_DATASET COMPOUND 4323
## 29            /Domain_10/STER/min_1      boom_2   H5I_GROUP              
## 30     /Domain_10/STER/min_1/boom_2 temperature H5I_DATASET COMPOUND 4323
## 31            /Domain_10/STER/min_1      boom_3   H5I_GROUP              
## 32     /Domain_10/STER/min_1/boom_3 temperature H5I_DATASET COMPOUND 4323
## 33                  /Domain_10/STER      min_30   H5I_GROUP              
## 34           /Domain_10/STER/min_30      boom_1   H5I_GROUP              
## 35    /Domain_10/STER/min_30/boom_1 temperature H5I_DATASET COMPOUND  147
## 36           /Domain_10/STER/min_30      boom_2   H5I_GROUP              
## 37    /Domain_10/STER/min_30/boom_2 temperature H5I_DATASET COMPOUND  147
## 38           /Domain_10/STER/min_30      boom_3   H5I_GROUP              
## 39    /Domain_10/STER/min_30/boom_3 temperature H5I_DATASET COMPOUND  147

```

```{r}
#now we can use this object to pull group paths from our file!
fiu_struct[3,1]

## [1] "/Domain_03/OSBS"

```

```{r}
## Let's view the metadata for the OSBS group
OSBS  <- h5readAttributes(f, fiu_struct[3,1])
#view the attributes
OSBS

## $LatLon
## [1] "29.68927/-81.99343"
## 
## $`Site Name`
## [1] "Ordway-Swisher Biological Station Site"

```

```{r}
# Grab the lat and long from the data
# note we might want to format the lat and long differently 
# This format is more difficult to extract from R!
OSBS$LatLon
```


## Part 2 - Workflows to Extract and Plot From Multiple Groups
The NEON HDF5 file that we are working with contains temperature data collected for three days (a very small subset of the available data) by one sensor. What if we wanted to create a plot that compared data across sensors or sites? To do this, we need to compare data stored within different nested groups within our H5 file.

### Data From Different Sensors Located at Different Levels, At One NEON Field Site

Let’s first compare data across temperature sensors located at one site. First, we’ll loop through the HDF5 file and build a new data frame that contains temperature data for each boom on the tower. We’ll use the 1-minute averaged data from the NEON field site: Ordway Swisher Biological Station located in Florida.

```{r}
#r compare temperature data for different booms at the Ordway Swisher site.
library(dplyr)
library(ggplot2)


# use dplyr to subset data by dataset name (temperature)
# and site / 1 minute average
newStruct <- fiu_struct %>% 
  filter(grepl("temperature", name), grepl("OSBS/min_1", group))

#create final paths to access each temperature dataset
paths <- paste(newStruct$group, newStruct$name, sep="/")


```

The above code uses dplyr to filter data. Let’s break the code down. Read more
about the dplyr package here

fiu_struct, defined above in the code, is the structure of our HDF5 file that we
returned using h5ls. grepl looks for a text pattern. Type help(grepl) to learn
more. If we want to return all “paths” in the HDF file that contain the word
temperature in the $name column, then we type grepl("temperature",name) filter
allows us to look for multiple strings in one command. help(filter) %>% is a
pipe - syntax specific to the dplyr package. It allows you to ‘chain’ or combine
multiple queries together into one, concise, line of code. Pulling this
together, type, fiu_struct %>% filter(grepl("OSBS/min_1",group)) in to the R
console. What happens?

Next, we will create a loop that will populate the final data.frame that
contains information for all booms in the site that we want to plot.

```{r}
#create a new, empty data.frame
ord_temp <- data.frame()


# loop through each temp dataset (all selected groups) and add to data.frame
for(i in paths){
  datasetName <- i
  print(datasetName) 
  #read in each dataset in the H5 list
  dat <- h5read(f, datasetName)            # read the contents in the dataset
  # add boom name to data.frame
  print(strsplit(i, "/")[[1]][5]) 
  dat$boom <- strsplit(i,"/")[[1]][5]      # add column with boom number
  ord_temp <- rbind(ord_temp, dat)         
}

## [1] "/Domain_03/OSBS/min_1/boom_1/temperature"
## [1] "boom_1"
## [1] "/Domain_03/OSBS/min_1/boom_2/temperature"
## [1] "boom_2"
## [1] "/Domain_03/OSBS/min_1/boom_3/temperature"
## [1] "boom_3"
## [1] "/Domain_03/OSBS/min_1/boom_5/temperature"
## [1] "boom_5"
## [1] "/Domain_03/OSBS/min_1/tower_top/temperature"
## [1] "tower_top"

head(ord_temp)
```


The loop above iterates through the file and grabs the temperature data for each boom in the 1 minute data series for Ordway. It also adds the boom name to the end of the data.frame as follows:

for i in path$path: loop through each path in the path object. NOTE: the boom 4 sensor was not operational when this HDF5 file was created, which is why there is no boom 4 in our list! Thus we will need do iterate through the data 5 times instead of 6.
dat <- h5read(f,i): read in the temperature dataset from our hdf5 file (f) for path i.
dat$boom <- strsplit(i,"/")[[1]][5]: add the boom name to a column called boom` in our data.frame
ord_temp <- rbind(ord_temp,dat): append dataset to the end of the data.frame


## Mini Challenge

Modify the loop above so that it adds both the boom name, the site name and the data type (1 minute) as columns in our data frame.

### Cleaning Up Dates

The dates field in our data frame aren’t imported by default in “date format”. We need to tell R to format the information as a date. Formatting out date fields also allows us to properly label the x axis of our plots.

Once the dates have been formatted we can create a plot with cleaner X axis labels.

```{r}
#fix the dates
ord_temp$date <- as.POSIXct(ord_temp$date,format = "%Y-%m-%d %H:%M:%S", tz = "EST")

# a nice R packages that helps with date formating is scale.
# install.packages("scales")
library("scales")
#plot the data
ggplot(ord_temp,aes(x=date,y=mean,group=boom,colour=boom))+
  geom_path()+
  ylab("Mean temperature") + xlab("Date")+
  theme_bw()+
  ggtitle("3 Days of temperature data at Ordway Swisher")+
  scale_x_datetime(breaks=pretty_breaks(n=4))
```


### Data from different sites

Next, let’s compare temperature at two different sites: Ordway Swisher Biological Station located in Florida and North Sterling located in Central Colorado. This time we’ll plot data averaged every 30 minutes instead of every minute. We’ll need to modify our search strings a bit. But we can still re-use most of the code that we just built.


```{r}
#grab just the paths to temperature data, 30 minute average
pathStrux <- fiu_struct %>% 
  filter(grepl("temperature",name), grepl("min_30",group)) 

#create final paths
paths <- paste(pathStrux$group, pathStrux$name, sep="/")

#create empty dataframe
temp_30 <- data.frame()

for(i in paths){
  #create columns for boom name and site name
  boom <-  strsplit(i,"/")[[1]][5]
  site <- strsplit(i,"/")[[1]][3]
  dat <- h5read(f,i)
  dat$boom <- boom
  dat$site <- site
 temp_30 <- rbind(temp_30,dat)
}

#Assign the date field to a "date" format in R
temp_30$date <- as.POSIXct(temp_30$date,format = "%Y-%m-%d %H:%M:%S")

# generate a mean temperature for every date across booms
temp30_sum <- temp_30 %>% 
  group_by(date, site) %>% 
  summarise(mean = mean(mean))

#Create plot!
ggplot(temp30_sum, aes(x=date,y=mean, group=site, colour=site)) + 
  geom_path()+ylab("Mean temperature, 30 Minute Average") + 
  xlab("Date")+
  theme_bw()+
  ggtitle("Comparison of Ordway-Swisher Biological Station (FL) vs North Sterling (CO)") +
  scale_x_datetime( breaks=pretty_breaks(n=4))
```

```{r}
H5close()
```

