---
title: "R Notebook"
output: html_notebook
---


Intorduction to HDF5

The following code examples create an HDF5 ﬁle and ﬁll it with some random data
to illustrate the functions we can use to interact with HDF5 ﬁles.

```{r}
#We use the rhdf5 package
library(rhdf5)
filename <- tempfile() #using a temporary file
h5createFile(filename) #creating the basic file layoutញ剆䕅淑
```


```{r}
h5ls(filename)
```

Let us use h5write to write some stuﬀ into the ﬁle (we use small examples, in a
nucleotide tally ﬁle the array dimensions will be several orders of magnitude
larger).

```{r}
# write an 4x5 array
h5write(obj = array(rnorm(20), dim=c(4,5)), file = filename, name = "/RandomArray")

# write a 100x1 vector
h5write(obj = rpois(100, 10), file = filename, name = "/PoissonVector")
h5ls(filename)
```

As you can see each dataset shows up in the listing of the ﬁle content
(h5ls(filename)). We can use h5read to extract the data from the ﬁle again, if
we simpy specify a ﬁlename and dataset name we will retrieve the whole dataset.

Since this is a really bad idea for large scale projects (e.g. nucleotide
tallies of human samples), we can also use subsetting to extract parts of the
data. Have a look at ?h5read if you want to know more about the possible ways of
retrieving data.￿

```{r}
h5read(file = filename, name = "/PoissonVector")
```


```{r}
# extract a part
h5read(file = filename, name = "/RandomArray", index = list(2:4, 3:5))
```

This should give you a basic idea of how we can interact with HDF5 ﬁles on a low
level. Luckily there are wrapper functions in h5vc that we can use to access the
ﬁle in a more eﬃcient way. Those will be discussed in the following sections.￿


