---
title: "HDF5 examples"
output: html_notebook
---


## 1st HDF5 example: Creating file and groups

```{r}
# Create an empty HDF5 file
library(rhdf5)
h5createFile("myhdf5file.h5")
```


The HDF5 file can contain a group hierarchy. We create a number of groups and list the file content afterwards.

```{r}
# create group foo
h5createGroup("myhdf5file.h5","foo")
```


```{r}
# create group baa
 h5createGroup("myhdf5file.h5","baa")
```


```{r}
# create sub group under foo
h5createGroup("myhdf5file.h5","foo/foobaa")
```


```{r}
# list what's in it
h5ls("myhdf5file.h5")
```



### 2nd example: writing and reading objects

```{r}
A = matrix(1:10, nr=5, nc=2)
h5write(A, "myhdf5file.h5","foo/A")

B = array(seq(0.1,2.0,by=0.1), dim = c(5, 2, 2))    # 5x2x2 array
attr(B, "scale") <- "liter"
h5write(B, "myhdf5file.h5","foo/B")

C = matrix(paste(LETTERS[1:10], LETTERS[11:20],     # matrix 2x5
                 collapse=""), nr=2,nc=5)
h5write(C, "myhdf5file.h5","foo/foobaa/C")

df = data.frame(1L:5L,seq(0,1,length.out=5),
c("ab","cde","fghi","a","s"), stringsAsFactors=FALSE)
h5write(df, "myhdf5file.h5","df")

h5ls("myhdf5file.h5")
```


```{r}
# create new dataset
h5createDataset("myhdf5file.h5", "foo/S", c(5,8), 
                storage.mode = "integer", chunk=c(5,1), level=7)
h5ls("myhdf5file.h5")
```

```{r}
h5write(matrix(1:5, nr=5, nc=1), file="myhdf5file.h5",         # matrix 5x1
         name="foo/S", index=list(NULL,1))                     # it will write only a vector of S
h5ls("myhdf5file.h5")
```

```{r}
# we can see here that only the first column of the matrix S was written
h5read("myhdf5file.h5", "foo/S")
```

```{r}
h5write(6:10, file="myhdf5file.h5",              # write the 1st row of S, columns 2 thru 6
        name="foo/S", index = list(1,2:6))

h5read("myhdf5file.h5", "foo/S")
```

```{r}
# write data in S on rows 1 thru 5 and columns 3 thru 8
h5write(matrix(11:40, nr=5, nc=6), file="myhdf5file.h5",
        name="foo/S", index=list(1:5, 3:8))

h5read("myhdf5file.h5", "foo/S")
```

```{r}
# write a sequence in rows 3,4 and columns 1,2
h5write(matrix(141:144, nr=2, nc=2), file="myhdf5file.h5", 
        name="foo/S", index=list(3:4, 1:2))

h5read("myhdf5file.h5", "foo/S")
```

```{r}
# write sequence 2x2 in rows 2, 3and columns 3 and 6 only
h5write(matrix(151:154,nr=2,nc=2), file="myhdf5file.h5", 
        name="foo/S", index=list(2:3, c(3,6)))
h5read("myhdf5file.h5", "foo/S")
```

```{r}
# read rows 2:3 and columns 2:3
h5read("myhdf5file.h5", "foo/S", index=list(2:3, 2:3))
```

```{r}
# read specific columns
h5read("myhdf5file.h5", "foo/S", index=list(2:3,c(2,4)))
```

```{r}
# read 4 specified columns
h5read("myhdf5file.h5", "foo/S", index=list(2:3,c(1,2,4,5)))
```

###

```{r}
# create another dataset H with 5x8 
h5createDataset("myhdf5file.h5", "foo/H", c(5,8), 
                storage.mode = "integer",
                chunk=c(5,1), level=7)
h5read("myhdf5file.h5", "foo/H")
```

```{r}
h5write(matrix(1:5,nr=5,nc=1), file="myhdf5file.h5", name="foo/H", 
        start=c(1,1))
h5read("myhdf5file.h5", "foo/H")
```


```{r}
h5write(6:10, file="myhdf5file.h5", name="foo/H",
        start=c(1,2), count=c(1,5))
h5read("myhdf5file.h5", "foo/H")
```

```{r}
h5write(matrix(11:40,nr=5,nc=6), file="myhdf5file.h5", name="foo/H",
        start=c(1,3))
h5read("myhdf5file.h5", "foo/H")
```


```{r}
h5write(matrix(141:144,nr=2,nc=2), file="myhdf5file.h5", name="foo/H", 
          start=c(3,1))
 h5read("myhdf5file.h5", "foo/H")
```

```{r}
# write small matrix in two different columns
h5write(matrix(151:154,nr=2,nc=2), file="myhdf5file.h5", name="foo/H",
        start=c(2,3), stride=c(1,3))

h5read("myhdf5file.h5", "foo/H")
```

```{r}
h5read("myhdf5file.h5", "foo/H", 
       start=c(2,1),             # start at row=2, col=1
       stride=c(1,3),            # only columns 1 and 3
       count=c(2,2),             # 
       block=c(1,2))
```


```{r}
h5ls("myhdf5file.h5", all=TRUE)
```


```{r}
h5ls("myhdf5file.h5", recursive=2)
```


-----------------------------------------------------------------

## R: Introduction to HDF5 Files in R
Source: http://neondataskills.org/HDF5/Intro-To-HDF5-In-R/

```{r}
# To access HDF5 files in R, we will use the rhdf5 library which is part of the 
#Bioconductor suite of R libraries.

#install rhdf5 package
#source("http://bioconductor.org/biocLite.R")
#biocLite("rhdf5")

#Call the R HDF5 Library
library("rhdf5")
```



```{r}
# Create hdf5 file
h5createFile("vegData.h5")

## [1] TRUE

h5createFile()
## Error in h5createFile(): argument "file" is missing, with no default


```

```{r}
#create a group called aNEONSite within the H5 file
h5createGroup("vegData.h5", "aNEONSite")

## [1] TRUE

#view the structure of the h5 we've created
h5ls("vegData.h5")

##   group      name     otype dclass dim
## 0     / aNEONSite H5I_GROUP
```

```{r}
# create some sample, numeric data 
a <- rnorm(n=40, m=1, sd=1) 
someData <- matrix(a,nrow=20,ncol=2)
```




```{r}
# add some sample data to the H5 file located in the aNEONSite group
# we'll call the dataset "temperature"
h5write(someData, file = "vegData.h5", name="aNEONSite/temperature")

# let's check out the H5 structure again
h5ls("vegData.h5")

##        group        name       otype dclass    dim
## 0          /   aNEONSite   H5I_GROUP              
## 1 /aNEONSite temperature H5I_DATASET  FLOAT 20 x 2
```


```{r}
# we can look at everything too 
# but be cautious using this command!
h5dump("vegData.h5")

## $aNEONSite
## $aNEONSite$temperature
##              [,1]        [,2]
##  [1,]  0.57002015  1.65021857
##  [2,] -0.11734323  0.67155051
##  [3,]  1.09386684  1.50329628
##  [4,]  0.67163631  0.30794922
##  [5,]  0.95862965 -0.78184862
##  [6,] -0.12896460  0.59268695
##  [7,]  0.05685202  2.28120086
##  [8,]  3.26047306  0.88577061
##  [9,] -0.30819170  0.47880539
## [10,]  0.96746933  1.82340641
## [11,] -0.54678513  2.22194159
## [12,]  1.29111825  1.05592524
## [13,]  0.16009552  0.41711324
## [14,]  1.05293700  0.07952516
## [15,]  1.92990794 -0.41059841
## [16,]  1.37178810  1.29995404
## [17,]  0.43722480  1.64796389
## [18,]  2.28480893  1.72972411
## [19,]  1.33004885 -0.61370131
## [20,]  1.99261923  1.64787132

#Close the file. This is good practice.
H5close()
```

```{r}
# Add Metadata (attributes)

# Let’s add some metadata (called attributes in HDF5 land) to our dummy temperature data. 
# First, open up the file.

#open the file, create a class
fid <- H5Fopen("vegData.h5")
#open up the dataset to add attributes to, as a class
did <- H5Dopen(fid, "aNEONSite/temperature")

# Provide the NAME and the ATTR (what the attribute says) 
# for the attribute.
h5writeAttribute(did, attr="Here is a description of the data",
                 name="Description")
h5writeAttribute(did, attr="Meters",
                 name="Units")

#let's add some attributes to the group
did2 <- H5Gopen(fid, "aNEONSite/")
h5writeAttribute(did2, attr="San Joaquin Experimental Range",
                 name="SiteName")
h5writeAttribute(did2, attr="Southern California",
                 name="Location")

#close the files, groups and the dataset when you're done writing to them!
H5Dclose(did)
H5Gclose(did2)
H5Fclose(fid)
```



### Working with an HDF5 File in R
Now that we’ve created our H5 file, let’s use it! First, let’s have a look at the attributes of the dataset and group in the file.

```{r}
#look at the attributes of the precip_data dataset
h5readAttributes(file = "vegData.h5", 
                 name = "aNEONSite/temperature")

## $Description
## [1] "Here is a description of the data"
## 
## $Units
## [1] "Meters"

#look at the attributes of the aNEONsite group
h5readAttributes(file = "vegData.h5", 
                 name = "aNEONSite")

## $Location
## [1] "Southern California"
## 
## $SiteName
## [1] "San Joaquin Experimental Range"

# let's grab some data from the H5 file
testSubset <- h5read(file = "vegData.h5", 
                 name = "aNEONSite/temperature")

testSubset2 <- h5read(file = "vegData.h5", 
                 name = "aNEONSite/temperature",
                 index=list(NULL,1))
H5close() 
# Once we've extracted data from our H5 file, we can work with it in R. 


#create a quick plot of the data
hist(testSubset2)
```


### Challenge –

Time to test your skills. Open up the D17_2013_SJER_vegStr.csv in R.

* Create a new HDF5 file called vegStructure.
* Add a group in your HDF5 file called SJER.
* Add the veg structure data to that folder.
* Add some attributes the SJER group and to the data.
* Now, repeat the above with the D17_2013_SOAP_vegStr csv.
* Name your second group SOAP

Some code is below to remind you how to import a CSV into R.



```{r}
veg <- read.csv("D17_2013_SOAP_vegStr.csv")
```



```{r}
# Create hdf5 file
h5createFile("vegStructure.h5")

#create a group called aNEONSite within the H5 file
h5createGroup("vegStructure.h5", "SJER")

# add some sample data to the H5 file located in the SJER  group
h5write(veg, file = "vegStructure.h5", name="SJER/veg")

# let's check out the H5 structure again
h5ls("vegStructure.h5")

```

```{r}
h5dump("vegStructure.h5")
```

```{r}
H5close() 
```


