---
title: "HDF5 - creating well files #01"
output: html_notebook
---

## Create the HDF5 file
```{r}
library(rhdf5)

wells.dir <- "./"

wellFile <- file.path(wells.dir, "WOODLIB.01.hdf5")
if (file.exists(wellFile)) file.remove(wellFile)

h5createFile(wellFile)
```

## add master groups

```{r}
fieldCase <- "GUO"            # field name
wellName  <- "Example_21"     # well name

group <- paste(fieldCase, wellName, sep = "/")
h5createGroup(wellFile, fieldCase)         # create the toplevel group first

h5createGroup(wellFile, group)             # add the well to the field case

# create sub-group based on current date and time
stamp <- format(Sys.time(), "%Y%m%d.%H%M")

wellEvent <- paste(group, stamp, sep = "/")

h5createGroup(wellFile, wellEvent) 
```



## add datasets

```{r}
# add a dataset with the well name
var.name <- "well.name"
var.path <- paste(group, var.name, sep = "/")

h5createDataset(wellFile, 
                var.path, 
                c(1,1), 
                storage.mode = "character", 
                size = 15)

```


### saving individual objects wit `h5save`
```{r}
well.name <- "EXAMPLE_01"
well.wh <- 100
well.bh <- 6000

h5save(well.name, well.bh, well.wh, file = wellFile)
h5ls(wellFile)
```

> with `h5save` we do not need to specify the name of the object



### saving a list that combines different data types

```{r}
H5Fopen(wellFile)

# a list is saved like a group

well.input <- list(oil.api = 40, 
                oil.visc = 5,
                gas.sg = 0.65,
                liq.rt = 1000,
                well.name = "well_01",
                segments = as.integer(30),
                deltaL = 1:40
                )

h5save(well.input, file = wellFile)
h5ls(wellFile)
H5close()
```



### saving with `h5write`
```{r}
H5Fopen(wellFile)

# a list is saved like a group
# each of the objects is a scalar
# the list or group has to be removed first in order to write it
# the list or group does not allow overwritting
# the list or group supports sequences. they are written as 1-col vectors


well.input2 <- list(oil.api = 40, 
                oil.visc = 5,
                gas.sg = 0.65,
                liq.rt = 1000,
                well.name = "well_01",
                segments = as.integer(50),
                deltaL = 1:50,
                water.sg = 1.01
                )

h5write(well.input2, file = wellFile, name = "well.input2")
h5ls(wellFile)
H5close()
```


### saving a list of scalars and a dataframe
```{r}
H5Fopen(wellFile)

# writing a dataframe
df3 = data.frame(long.int = 1L:5L, 
                sequence = seq(0,1, length.out=5), 
                strings = c("ab","cde","fghi","a","s"), 
                stringsAsFactors=FALSE)

well.input3 <- list(oil.api = 40, 
                oil.visc = 5,
                gas.sg = 0.65,
                liq.rt = 1000,
                well.name = "well_01",
                segments = as.integer(50),
                deltaL = 1:50,
                water.sg = 1.01,
                dataframe = df3
                )

h5write(well.input3, file = wellFile, name = "well.input3")
h5ls(wellFile)
H5close()
```

### writting attributes to group or list with `h5write`
```{r}
H5Fopen(wellFile)

# writing a dataframe
df3 = data.frame(long.int = 1L:5L, 
                sequence = seq(0,1, length.out=5), 
                strings = c("ab","cde","fghi","a","s"), 
                stringsAsFactors=FALSE)

well.input4 <- list(oil.api = 40, 
                oil.visc = 5,
                gas.sg = 0.65,
                liq.rt = 1000,
                well.name = "well_01",
                segments = as.integer(50),
                deltaL = 1:50,
                water.sg = 1.01,
                dataframe = df3
                )

h5write(well.input4, 
        file = wellFile, 
        name = "well.input4",
        write.attributes = TRUE
        )

h5ls(wellFile)
H5close()
```

> When write.attributes is TRUE, the group shows a red "A" letter.
The default attributes of the list are shown as name = "names" and values are the names of the members of the list.



## write attributes

```{r}
# Open a hdf5 file
fid <- H5Fopen(wellFile)

#open up the dataset to add attributes to, as a class
did <- H5Dopen(fid, "/well.input3/deltaL")           # did refers to the dataset

h5writeAttribute(did, 
                 attr = "feet",
                 name = "Units")

h5ls(wellFile)
H5close()
```


```{r}
# Open a hdf5 file
fid <- H5Fopen(wellFile)

#open up the dataset to add attributes to, as a class
did <- H5Dopen(fid, "/well.input3/gas.sg")           # did refers to the dataset

h5writeAttribute(did, 
                 attr = "specific gravity",
                 name = "Units")

did <- H5Dopen(fid, "/well.input3/liq.rt")           # did refers to the dataset
h5writeAttribute(did, 
                 attr = "barrel per day",
                 name = "Units")

h5ls(wellFile)
H5close()
```









































```{r}
# writing scalars
var.names <- read.csv("./extdata/variable_names.csv")
nRow <- nrow(var.names)

for (v in seq_len(nRow)) {
  #cat(v, var.names[v, 3], "\n")
  row <- as.character(var.names[v, 1])
  #print(unlist(row["var.name"]))
  cat(row, "\n")
  
}

```





```{r}
h5ls(wellFile)
```

