---
title: "How to append rows to an R data frame"
output: html_notebook
---

Source: http://stackoverflow.com/questions/20689650/how-to-append-rows-to-an-r-data-frame/38052208

```{r}
# use rbind
f1 <- function(n){
  df <- data.frame(x = numeric(), y = character())
  for(i in 1:n){
    df <- rbind(df, data.frame(x = i, y = toString(i)))
  }
  df
}

# use list
f2 <- function(n){
  df <- data.frame(x = numeric(), y = character(), stringsAsFactors = FALSE)
  for(i in 1:n){
    df[i,] <- list(i, toString(i))
  }
  df
}

# pre-allocate space
f3 <- function(n){
  df <- data.frame(x = numeric(n), y = character(n), stringsAsFactors = FALSE)
  for(i in 1:n){
    df$x[i] <- i
    df$y[i] <- toString(i)
  }
  df
}


# Use preallocated vectors
f4 <- function(n) {
  x <- numeric(n)
  y <- character(n)
  for (i in 1:n) {
    x[i] <- i
    y[i] <- i
  }
  data.frame(x, y, stringsAsFactors=FALSE)
}

library(microbenchmark)
mb <- microbenchmark(f1(1000), f2(1000), f3(1000), f4(1000), times = 10)
autoplot(mb)

```

The best solution is to pre-allocate space (as intended in R). The next-best solution is to use list, and the worst solution (at least based on these timing results) appears to be rbind.



## create, append, access
Suppose you simply don't know the size of the data.frame in advance. It can well be a few rows, or a few millions. You need to have some sort of container, that grows dynamically. Taking in consideration my experience and all related answers in SO I come with 4 distinct solutions:

rbindlist to the data.frame
Use data.table's fast set operation and couple it with manually doubling the table when needed.
Use RSQLite and append to the table held in memory.
data.frame's own ability to grow and use custom environment (which has reference semantics) to store the data.frame so it will not be copied on return.
Here is a test of all the methods for both small and large number of appended rows. Each method has 3 functions associated with it:

create(first_element) that returns the appropriate backing object with first_element put in.
append(object, element) that appends the element to the end of the table (represented by object).
access(object) gets the data.frame with all the inserted elements.

```{r}
# rbindlist to the data.frame
create.1<-function(elems)
{
  return(as.data.table(elems))
}

append.1<-function(dt, elems)
{ 
  return(rbindlist(list(dt,  elems),use.names = TRUE))
}

access.1<-function(dt)
{
  return(dt)
}


# data.table::set + manually doubling the table when needed
create.2<-function(elems)
{
  return(as.data.table(elems))
}

append.2<-function(dt, elems)
{
  n<-attr(dt, 'rowcount')
  if (is.null(n))
    n<-nrow(dt)
  if (n==nrow(dt))
  {
    tmp<-elems[1]
    tmp[[1]]<-rep(NA,n)
    dt<-rbindlist(list(dt, tmp), fill=TRUE, use.names=TRUE)
    setattr(dt,'rowcount', n)
  }
  pos<-as.integer(match(names(elems), colnames(dt)))
  for (j in seq_along(pos))
  {
    set(dt, i=as.integer(n+1), pos[[j]], elems[[j]])
  }
  setattr(dt,'rowcount',n+1)
  return(dt)
}

access.2<-function(elems)
{
  n<-attr(elems, 'rowcount')
  return(as.data.table(elems[1:n,]))
}



# SQL should be optimized for fast record insertion, so I initially had high hopes for RSQLite solution
create.3<-function(elems)
{
  con <- RSQLite::dbConnect(RSQLite::SQLite(), ":memory:")
  RSQLite::dbWriteTable(con, 't', as.data.frame(elems))
  return(con)
}

append.3<-function(con, elems)
{ 
  RSQLite::dbWriteTable(con, 't', as.data.frame(elems), append=TRUE)
  return(con)
}

access.3<-function(con)
{
  return(RSQLite::dbReadTable(con, "t", row.names=NULL))
}

# data.frame's own row-appending + custom environment.
create.4<-function(elems)
{
  env<-new.env()
  env$dt<-as.data.frame(elems)
  return(env)
}

append.4<-function(env, elems)
{ 
  env$dt[nrow(env$dt)+1,]<-elems
  return(env)
}

access.4<-function(env)
{
  return(env$dt)
}
```



```{r}
test <- function(id, n = 100)
{
  n <- n-1
  el <- list(a=1, b=2, c=3, d=4)                  # the row data
  o <- do.call(paste0('create.',id), list(el))    # create the row
  s <- paste0('append.',id)                     
  for (i in 1:n)
  {
    o <- do.call(s, list(o,el))                    # add the row
  }
  return(do.call(paste0('access.', id), list(o)))  # read the row
}
```


## The test suite:

For convenience I will use one test function to cover them all with indirect calling. (I checked: using do.call instead of calling the functions directly doesn't makes the code run measurable longer).
```{r}
library(microbenchmark)

rows <- c(10, 50, 100, 1000)
timings <- vector("list", length(rows))

i <- 1
for (tm in rows) {
  cat(tm, "\n")
  ti <- microbenchmark(test(1, n = tm),
                       test(2, n = tm),
                       test(3, n = tm), 
                       test(4, n = tm), times = 10)
  ti <- data.frame(ti, stringsAsFactors = FALSE)
  ti$nRow <- tm
  timings[[i]] <- ti
  i <- i + 1
  #autoplot(ti)  
  # print(ti)
}

timings <- data.table::rbindlist(timings)

source("plotBenchmarks.R")
print(plotTimings(timings))
```

### We exclude now test #3 and increase the # of rows


```{r}
library(microbenchmark)

rows <- c(10, 50, 100, 1000, 2000, 5000, 10000)
timings <- vector("list", length(rows))

i <- 1
for (tm in rows) {
  cat(tm, "\n")
  ti <- microbenchmark(test(1, n = tm),
                       test(2, n = tm),
                       test(4, n = tm), times = 10)
  ti <- data.frame(ti, stringsAsFactors = FALSE)
  cat(i, tm, dim(ti), "\n")
  ti$nRow <- tm
  timings[[i]] <- ti
  i <- i + 1
}

timings <- data.table::rbindlist(timings)

source("plotBenchmarks.R")
print(plotTimings(timings))
```




```{r}
extendDf <- function (df, n) {
    withFactors <- sum(sapply (df, function(X) (is.factor(X)) )) > 0
    nr          <- nrow (df)
    colNames    <- names(df)
    for (c in 1:length(colNames)) {
        if (is.factor(df[,c])) {
            col         <- vector (mode='character', length = nr+n) 
            col[1:nr]   <- as.character(df[,c])
            col[(nr+1):(n+nr)]<- rep(col[1], n)  # to avoid extra levels
            col         <- as.factor(col)
        } else {
            col         <- vector (mode=mode(df[1,c]), length = nr+n)
            class(col)  <- class (df[1,c])
            col[1:nr]   <- df[,c] 
        }
        if (c==1) {
            newDf       <- data.frame (col ,stringsAsFactors=withFactors)
        } else {
            newDf[,c]   <- col 
        }
    }
    names(newDf) <- colNames
    newDf
}
```


The function extendDf() extends a data frame with n rows.

As an example:
```{r}
aDf <- data.frame (l=TRUE, i=1L, n=1, c='a', t=Sys.time(), stringsAsFactors = TRUE)
extendDf (aDf, 2)
```

