---
title: "R Notebook"
output: html_notebook
---

## Using sdcMicro
Statistical Disclosure Control (SDC) for the generation of protected microdata for researchers and for public use.

This package includes all methods of the popular software mu-Argus plus several new methods. In comparison with mu-Argus the advantages of this package are that the results are fully reproducible even with the included GUI, that the package can be used in batch-mode from other software, that the functions can be used in a very flexible way, that everybody could look at the source code and that there are no time-consuming meta-data management is necessary. However, the user should have a detailed knowledge about SDC when applying the methods on data.

```{r}
# load the sdcMicro package
library(sdcMicro)

## select another keyVar and run localSupp once again,
#if you think the table is not fully protected
data(free1)
free1 <- as.data.frame(free1)

f <- freqCalc(free1, keyVars=1:3, w=30)
ind <- indivRisk(f)
## and now you can use the interactive plot for individual risk objects:
## plot(ind)
ind
```



```{r}
# load the sdcMicro package
library(sdcMicro)

## Data from mu-Argus:
## Global recoding:
data(free1)
free1 <- as.data.frame(free1)

free1[, "AGE2"] <- globalRecode(free1[,"AGE"], c(1,9,19,29,39,49,59,69,100), labels=1:8)


```

```{r}
free1[, c("AGE", "AGE2")]
```



```{r}
data(free1)
free1 <- as.data.frame(free1)

## application to a vector
head(globalRecode(free1$AGE, breaks=c(1,9,19,29,39,49,59,69,100), labels=1:8))
table(globalRecode(free1$AGE, breaks=c(1,9,19,29,39,49,59,69,100), labels=1:8))
```

```{r}
## Fast generation of synthetic data with aprox.
# the same covariance matrix as the original one.
data(mtcars)
cov(mtcars[, 4:6])                     # original data
cov(dataGen(mtcars[, 4:6],n=200))      #synth data

pairs(mtcars[,4:6])
pairs(dataGen(mtcars[,4:6],n=200))
```

## addNoise
method: choose between ‘additive’, ‘correlated’, ‘correlated2’, ‘restr’, ‘ROMM’, ‘outdect’

```{r}
data(Tarragona)
a1 <- addNoise(Tarragona)     # creates a list with modified values (perturbed data)
names(a1)

head(a1$x)
head(as.data.frame(a1$xm))
```

```{r}
data(Tarragona)
a1 <- addNoise(Tarragona, method = "correlated")    # modified values (perturbed data)
names(a1)

head(a1$x)
head(as.data.frame(a1$xm))
```
