---
title: "R Notebook"
output: html_notebook
---


```{r}
library(rJava)
ospClasses <- "./inst/osp" 
.jinit()          # initialize Java-VM
.jaddClassPath(ospClasses)
.jclassPath()
```


```{r}
dp <- J("org.opensourcephysics.display.DrawingPanel")
df <- J("org.opensourcephysics.display.DrawingFrame")
```

    import org.opensourcephysics.display2d.ArrayData;
    DrawingPanel drawingPanel = new DrawingPanel();
    drawingPanel.setShowCoordinates(false);
    DrawingFrame frame = new DrawingFrame(drawingPanel);
    ArrayData arraydata = new ArrayData(32, 32, 1);
    arraydata.setScale(-1, 1, -1, 1);
        double[][] data = arraydata.getData()[0];
    double x = arraydata.getLeft(), dx = arraydata.getDx();

    
```{r}
dpanel <- new(J("org.opensourcephysics.display.DrawingPanel"))
dpanel$setShowCoordinates(FALSE)
dframe <- new(J("org.opensourcephysics.display.DrawingFrame"))
```


```{r}
arraydata <- new(J("org.opensourcephysics.display2d.ArrayData"), 
                 as.integer(32), as.integer(32), as.integer(1))
arraydata$setScale(-1, 1, -1, 1)

adata <- arraydata$getData()
# data$length
x <- arraydata$getLeft()
dx <- arraydata$getDx()
```


        for(int i = 0, nx = data.length;i<nx;i++) {
          double y = arraydata.getTop(), dy = arraydata.getDy();
          for(int j = 0, ny = data[0].length;j<ny;j++) {
            data[i][j] = y*x; // magnitude
            y += dy;
          }
          x += dx;
        }


```{r}
head(.jevalArray(adata, simplify = TRUE))
```


```{r}
for (i in seq(1, 32, 1 )) {
    y <- arraydata$getTop()
    dy <- arraydata$getDy()
    for (j in seq(1, 32, 1)) {
        adata[[i,j]] <- y * x
        y <- y + dy
    }
    x <- x + dx
}
```


