---
title: "R Notebook"
output: html_notebook
---

Examples. We load the marioKart data object from the openintro package, which provides information about 143 auctions for the Wii game, Mario Kart. 

* (0) A new data set with two outliers is returned and saved to mk0. 

* (1) A plot is constructed for the number of steering wheels included in the auction against the auction’s total price. Notice that the variables are accessed in the expressions without explicitly invoking the mk0 data frame. 

* (2) An object is removed, so it will not be included in the returned value from within. 

* (3) Other values can also be modified for the object returned from within. 

* (4) Lastly, variables may be added to the outgoing object for within, as is done here with the end price of the auction.


```{r}

# install.packages("openintro")

library(openintro)

data(marioKart)
names(marioKart)
dim(marioKart)

#______ 0. Removing Two Outliers ______#
mk0 <- marioKart[marioKart$totalPr < 100,]


#______ 1. Create Plot ______#
cat("#______ 1. Create Plot ______#\n")
png("wheels_totalPr.png")
with(mk0, {
           boxplot(totalPr ~ wheels)
           points(wheels+1.1, totalPr, col=4)
          })
# dev.off()


#______ 2. Remove One Column ______#
cat("#______ 2. Remove One Column ______#\n")
mk2 <- within(mk0, rm(title))
names(mk2)


#______ 3. Change Values ______#
cat("#______ 3. Change Values ______#\n")
mk0$totalPr[50]
mk0$startPr[25]
mk3 <- within(mk0, { # Would not typically do...
                     # this is just an example
                    totalPr[50] <- 88.59
                    startPr[25] <- 85.00
                   })
mk3$totalPr[50]
mk3$startPr[25]


#______ 4. Constuct Auction End Price ______#
cat("#______ 4. Constuct Auction End Price ______#")
mk4 <- within(mk0, endPrice <- totalPr - shipPr)
all.equal(mk4$totalPr - mk4$shipPr, mk4$endPrice)
names(mk4)
```


```{r}
# in this case, we don't need anorexia as a prefix
library(MASS)
with(anorexia, {
    anorex.1 <- glm(Postwt ~ Prewt + Treat + offset(Prewt),
                    family = gaussian)
    summary(anorex.1)
})
```


```{r}
# in this example we modify the data frame and write the modifications
# in another data frame

aq <- within(airquality, {     # Notice that multiple vars can be changed
        lOzone <- log(Ozone)                  # calculate log
        Month  <- factor(month.abb[Month])    # convert to factor
        cTemp  <- round((Temp - 32) * 5/9, 1) # From Fahrenheit to Celsius
        S.cT   <- Solar.R / cTemp             # using newly created variable
        rm(Day, Temp)                         # remove 2 variables
})

head(aq)   # new data frame
```


```{r}
# on the other hand, `with` doesn't do antything
aq <- with(airquality, {     # Notice that multiple vars can be changed
    lOzone <- log(Ozone)
    Month <- factor(month.abb[Month])
    cTemp <- round((Temp - 32) * 5/9, 1) # From Fahrenheit to Celsius
    S.cT <- Solar.R / cTemp  # using the newly created variable
    rm(Day, Temp)
})
head(aq)
```

http://stackoverflow.com/questions/21827572/what-is-the-difference-between-with-and-within-in-r

```{r}
require(plyr)

# with(baseball, baseball[id=="ansonca01",])
with(baseball, baseball[id=="ansonca01",])
```


```{r}
within(baseball, baseball[id=="ansonca01", ])
# code doesn’t modify anything inside baseball, the unmodified baseball is returned. with on the other hand doesn’t return the object, it returns expr.
```


```{r}
head(within(cars, speed[dist < 20] <- 1))
```



http://stackoverflow.com/questions/21827572/what-is-the-difference-between-with-and-within-in-r

```{r}
df <- data.frame(a=1:5,b=2:6)

with(df, 
     {c <- a + b
     df})

within(df, 
       {c <- a + b
       df} )
```

