---
title: "R Notebook"
output: html_notebook
---

https://jamesmccaffrey.wordpress.com/2015/05/17/r-language-oop-using-r6/

The R language is used mostly by statisticians and data people. There are at least four, quite different, ways to write object oriented code in R — S3, S4, R5, R6. By far my favorite approach is to use the relatively new R6 version 2 (released October 2014).

Compared to the alternatives, the R6 OOP approach and syntax is closest to OOP in languages like C# and Python. Here’s a quick example. First, I created a text file that defines a Car class and saved it as CarClass.R on my machine:

```{r}
# CarClass.R

require("R6")

Car <- R6Class("Car",
  public = list(
    make = NULL,
    price = NULL,
    initialize = function(ma, pr) {
      self$make <- ma
      self$price <- pr
    },
    
    setMake = function(ma) { 
        self$make <- ma 
        },
    # setPrice = function(pr) { self$price <- pr },
    
    display = function() {
      cat("Make = ", self$make,
        " Price = ", self$price, "\n")
    }
  )
)
```



```{r}
someCar = Car$new("Audi", 40000)
someCar$display()
```

```{r}
# display using built-in print()
print(someCar)
```

Objects created using R6 get a built-in print() capability. To modify the Car object:

```{r}
someCar$setMake("BMW")
someCar$price  = 50000
someCar$display()
```

I modified the Car make field using the set-function, but modified the public price field directly. R6 objects are reference objects so you have to be careful when assigning one object to another. My bottom line: R6, very nice.