---
title: "R Notebook"
output: html_notebook
---

Source: http://stackoverflow.com/a/39929961/5270873

We have this Java class. Use `rJava` to call the methods after creating a
package JAR from the class.

        package hello;
        
            public class Hello extends Object {
        
                public String sayHello2(String name) {
                    String result = new String("Hello " + name);
                    return result;
                }
        
                public Hello sayHello3(String name) {
                    String result = new String("Hello " + name);
                    return new Hello();
                }
        
            } 
            
```{r}
# compile Java file. Same name as the class
system("javac -verbose ./inst/java/Hello.java")
```

To create a jar :

       java -cp .   hello/Hello.java
       jar cvf Hello.jar hello/Hello.class
       
       
```{r}
# function contributed by Bazz: stackoverflow.com/a/30802765/5270873 
command <- function(command, intern = TRUE, wait = FALSE){
  system(paste("cmd.exe /c", command), intern = T, wait = wait)
}

cmd <- paste("cd ./inst/java",                      # change to Java folder
             "javac hello/Hello.java",              # compile Java code
             "jar cvf Hello.jar hello/Hello.class", # create JAR
             sep = "&& ")

command(cmd)
```

```{r}
library(rJava)

.jinit()
.jaddClassPath(dir( "./inst/java", full.names=TRUE ))
.jclassPath()
hjw <- .jnew("hello.Hello")     # create instance of hell/Hello class
```

```{r}
outRef <- .jcall(hjw, "S", "sayHello2", "toto", evalString = FALSE)
.jstrVal(outRef)
```

```{r}
# And a call to a function returning a java reference:
outRef2 <- .jcall(hjw, "Lhello/Hello;", "sayHello3", "Universe", evalString = T)
.jstrVal(outRef2)

outRef3 <- .jcall(outRef2, "S", "sayHello2", "New Universe", evalString = FALSE)
.jstrVal(outRef3)
```




## R and System calls

Source: http://stackoverflow.com/a/30802765/5270873

```{r}
command <- function(command, intern = TRUE, wait = FALSE){
  system(paste("cmd.exe /c", command), intern = T, wait = wait)
}

cmdStr <- paste("cd", "./inst/java", "&&", "dir > aList.txt")
command("cd ./inst/java && dir > any.txt")
```

```{r}
# R and System calls
cmdStr <- paste("cd", "./inst/java", "&&", "dir > aList.txt", sep = " ")
command(cmdStr)
```

