---
title: "Open a Java Dialog"
output: html_notebook
---

Source: http://www.deducer.org/pmwiki/pmwiki.php?n=Main.Development#wwjoir

## Getting Started

Lets start by making a `JDialog` object. `JDialog` is a class used to make a
dialog window, and is a part of the Swing GUI library.

The first thing we need to do is create a variable representing the class. This
can be done with the `J` function, which takes as an argument the class location,
and returns a reference to that location (an S4 object of class `jclassName`).
`JDialog` is located in the javax.swing Java package, so we simply need to call.

```{r}
library(rJava)

Jdialog <- J("javax.swing.JDialog")
print(Jdialog)
```

Now that we have a reference to the JDialog class, we can a new object by calling the new function. new takes as its first argument a jclassName (e.g. JDialog), and any further arguments are passed to the JDialog constructor.

```{r}
myDialog <- new(Jdialog)
```

`myDialog` is now a reference (of S4 class jobjRef) to an instance of JDialog. We
don't see anything yet, because we have not made the dialog visible. There is a
JDialog method called setVisible which we can use to make the dialog visible.
Set visible takes a boolean (true/false) as a parameter. rJava transparently
takes care of the conversion between R logical and Java boolean. We will go into
conversions later on.

```{r}
myDialog$setVisible(TRUE)
```

You should now see a small empty dialog window.


```{r}
.jfloat(1)
```


```{r}
# careful! You may have to terminate R to close the Java window

.jcall("java/lang/System","S","getProperty","os.name")
if (!nzchar(Sys.getenv("NOAWT"))) {
  f <- .jnew("java/awt/Frame","Hello")
  .jcall(f,,"setVisible",TRUE)
}

```

