---
title: "R Notebook"
output: html_notebook
---


## 2.4. Inheritance
A class’ inheritance hiearchy can be specified when the class is defined using
the contains option. The contains option is a vector that lists the classes the
new class inherits from. In the following example we build on the Agent class
defined in the previous section. The idea is that we need agents that represent
a predator and two prey. We will focus on two predators for this example.

The hierarchy for the classes is shown in Figure 2.. In this example we have one
Prey class that is derived from the Agent class. There are two predator classes,
Bobcat and Lynx. The Bobcat class is derived from the Agent class, and the Lynx
class is derived from the Bobcat class. We will keep this very simple, and the
only methods associated with the new classes is a move method. For our purposes
it will only print out a message and set the values of the position and velocity
to demonstrate the order of execution of the methods associated with the
classes.

The first step is to create the three new classes.

```{r}
######################################################################
# Create the Prey class
#
# This is used to represent a prey animal
Prey <- setClass(
        # Set the name for the class
        "Prey",

        # Define the slots - in this case it is empty...
        slots = character(0),

        # Set the default values for the slots. (optional)
        prototype = list(),

        # Make a function that can test to see if the data is consistent.
        # This is not called if you have an initialize function defined!
        validity = function(object) {
                if(sum(object@velocity^2)>70.0) {
                        return("The velocity level is out of bounds.")
                }
                return(TRUE)
        },

        # Set the inheritance for this class
        contains = "Agent"
        )



######################################################################
# Create the Bobcat class
#
# This is used to represent a smaller predator
Bobcat <- setClass(
        # Set the name for the class
        "Bobcat",

        # Define the slots - in this case it is empty...
        slots = character(0),

        # Set the default values for the slots. (optional)
        prototype=list(),

        # Make a function that can test to see if the data is consistent.
        # This is not called if you have an initialize function defined!
        validity=function(object) {
                if(sum(object@velocity^2)>85.0) {
                        return("The velocity level is out of bounds.")
                }
                return(TRUE)
        },

        # Set the inheritance for this class
        contains = "Agent"
        )

######################################################################
# Create the Lynx class
#
# This is used to represent a larger predator
Lynx <- setClass(
        # Set the name for the class
        "Lynx",

        # Define the slots - in this case it is empty...
        slots = character(0),

        # Set the default values for the slots. (optional)
        prototype = list(),

        # Make a function that can test to see if the data is consistent.
        # This is not called if you have an initialize function defined!
        validity = function(object) {
                if(sum(object@velocity^2)>95.0) {
                        return("The velocity level is out of bounds.")
                }
                return(TRUE)
        },

        # Set the inheritance for this class
        contains = "Bobcat"
        )
```

The inheritance is specified using the contains option in the setClass command.
Note that this can be a vector allowing for multiple inheritance. We choose not
to use that to keep things simpler. If you are feeling like you need more
self-loathing in your life you should try it out and experiment.

Next we define a method, move, for the new classes. We will include methods for
the Agent, Prey, Bobcat, and Lynx classes. The methods do not really do anything
but are used to demonstrate the idea of how methods are executed.

```{r}
# create a method to move the agent.
setGeneric(name="move", def = function(theObject) {
                               standardGeneric("move")
                       })

setMethod(f="move",
                      signature="Agent",
                      definition=function(theObject)
                      {
                              print("Move this Agent dude")
                              theObject <- setVelocity(theObject,c(1,2))
                              validObject(theObject)
                              return(theObject)
                      }
                      )

setMethod(f="move",
                      signature="Prey",
                      definition=function(theObject)
                      {
                              print("Check this Prey before moving this dude")
                              theObject <- callNextMethod(theObject)
                              print("Move this Prey dude")
                              validObject(theObject)
                              return(theObject)
                      }
                      )

setMethod(f="move",
                      signature="Bobcat",
                      definition=function(theObject)
                      {
                              print("Check this Bobcat before moving this dude")
                              theObject <- setLocation(theObject,c(2,3))
                              theObject <- callNextMethod(theObject)
                              print("Move this Bobcat dude")
                              validObject(theObject)
                              return(theObject)
                      }
                      )

setMethod(f="move",
                      signature="Lynx",
                      definition=function(theObject)
                      {
                              print("Check this Lynx before moving this dude")
                              theObject <- setActive(theObject,FALSE)
                                theObject <- callNextMethod(theObject)
                              print("Move this Lynx dude")
                              validObject(theObject)
                              return(theObject)
                      }
                      )
```

There are a number of things to note. First each method calls the `callNextMethod`
command. This command will execute the next version of the same method for the
previous class in the hierarchy. Note that I have included the arguments (in the
same order) as those called by the original function. Also note that the
function returns a copy of the object and is used to update the object passed to
the original function.

Another thing to note is that the methods associated with the Lync, Bobcat, and
Agent classes arbitrarily change the values of the position, velocity, and
activity for the given object. This is done to demonstrate the changes that take
place and reinforce the necessity for using the callNextMethod function the way
it is used here.

Finally, it should be noted that the `validObject` command is called in every
method. You should try adding a print statement in the validity function. You
might find that the order is a bit odd. You should experiment with this and play
with it. There are times you do not get the expected results so be careful!

We now give a brief example to demonstrate the order that the functions are
called. In the example we create a Bobcat object and then call the move method.
We next create a Lynx object and do the same. We print out the slots for both
agents just to demonstrate the values that are changed.

```{r}
robert <- Bobcat()
robert

robert <- move(robert)
robert

lionel <- Lynx()
lionel

lionel <- move(lionel)
lionel
```

