---
title: "R Notebook"
output: html_notebook
---

## R fundamentals
```{r}
x = 10
y = x

ls()
# [1] "x" "y"
```


```{r}
# variable names that do not follow the standard rules have to be quoted or
# use backticks
`_foo` = 10
"10:10" = 20

ls()

# [1] "_foo"  "10:10" "x"     "y" 
```

## attributes
```{r}
x = 1:10
attr(x, "foo") = 11   # add an attribute to x
x
#  [1]  1  2  3  4  5  6  7  8  9 10
# attr(,"foo")
# [1] 11
```

## intro to OOP in R

```{r}
mean
# function (x, ...) 
# UseMethod("mean")
# <bytecode: 0x0000000009d70470>
# <environment: namespace:base>
```

We can see all the deﬁned methods for this function using the methods command.

```{r}
methods(mean)
```

And see that they all begin with the name mean, then a period. 

## Numerical computing
One of the strengths of R is its various numerical computing capabilities. It
is important to remember that computers cannot represent all numbers and
that machine computation is not identical to computation with real numbers.

And a classical example of the problem is given in the code below.

```{r}
a = sqrt(2)
a * a == 2
# [1] FALSE

a * a - 2
# [1] 4.440892e-16
```

      
      > .Machine
      
      $double.eps
      [1] 2.220446e-16
      
      $double.neg.eps
      [1] 1.110223e-16
      
      $double.xmin
      [1] 2.225074e-308
      
      $double.xmax
      [1] 1.797693e+308
      
      $double.base
      [1] 2
      
      $double.digits
      [1] 53
      
      $double.rounding
      [1] 5
      
      $double.guard
      [1] 0
      
      $double.ulp.digits
      [1] -52
      
      $double.neg.ulp.digits
      [1] -53
      
      $double.exponent
      [1] 11
      
      $double.min.exp
      [1] -1022
      
      $double.max.exp
      [1] 1024
      
      $integer.max
      [1] 2147483647
      
      $sizeof.long
      [1] 4
      
      $sizeof.longlong
      [1] 8
      
      $sizeof.longdouble
      [1] 16
      
      $sizeof.pointer
      [1] 8


## Environments
 In the code below, we create
an environment, create the symbol value pair that relates the symbol a to the
value 10 and then list the contents of the hash table.

```{r}
e1 = new.env(hash = TRUE)
e1$a = 10
ls(e1)
# [1] "a"
```

```{r}
e1[["a"]]
# [1] 10
```

```{r}
e1$b = 20
e1$c = 30

```

In order to retrieve multiple values simultaneously from an environment, the
mget function should be used.

```{r}
mget(c("a", "b", "c"), e1)
```

In the next code segment, an environment, e1, is created and has some
values assigned in to it. Then a function is deﬁned and that function has
some free variables (variables that are not parameters and are not deﬁned
in the function). We then make e1 be the environment associated with the
function and then the free variables will obtain values from e1. Then we
change the value of one of the free variables by accessing e1 and that changes
the behavior of the function, which demonstrates that no copy of e1 was made.

```{r}
e1 = new.env()
e1$z = 10

f = function(x) {
  x + z
}

environment(f) = e1
f(10)

e1$z = 20
f(10)
```

Next, we demonstrate the semantics of rm in this context. If we remove e1,
what should happen to f? If the eﬀect of the command environment(f) = e1
was to make a copy of e1, then rm(e1) should have no eﬀect, but we know
that no copy was made and yet, as we see, removing e1 appears to have no
eﬀect.
```{r}
rm(e1)
f(10)
```

What rm(e1) does is to remove the binding between the symbol e1 and the
internal data structure that contains the data, but that internal data structure
is itself left alone. Since it can also be reached as the environment of f, it will remain available.


## Managing your R session

```{r}
R.version$platform

Sys.info()

.Platform

capabilities()
```

## 2.4 Language basics

```{r}
colSums
```

```{r}
get("+")

x = 1:4
myP = get("+")

myP(x, 5)
```


users can deﬁne their
own using any text string in place of any. The function should be a function of
two arguments, although currently this is not checked. In the example below
we deﬁne a simple operator that pastes together its two arguments.

```{r}
"%p%" = function(x, y) paste(x, y, sep = "")

"hi" %p% "there"
```


```{r}
foo = function(x) {
  if (x < 3)
    list() + x
  else {
    if (x < 10)
      warning("ouch")
    else 33
  }
}

tryCatch(foo(2), error = function(e) "an error", 
         warning = function(e) "a warning")
```

```{r}
tryCatch(foo(5), error = function(e) "an error", 
         warning = function(e) "a warning")
```

## 2.11 Exception handling
```{r}
tryCatch(foo(29))
```


```{r}
even = function(i) i %% 2 == 0

testEven = function(i) if (even(i) ) i else stop("not even")

vals = NULL

withCallingHandlers({ 
  for (i in seq_len(10)) { 
    val = withRestarts(testEven(i), 
                       skipError=function() return(NULL))
    if (!is.null(val))
      vals = c(vals, val) }
  }, 
  error=function(e) invokeRestart("skipError"))

vals
```

## 2.12.4 Indirect function invocation

```{r}
b = get("foo")
b(23)
# [1] 33
```

```{r}
do.call("foo", list(23))
```

```{r}
x = expression(1:10)
x
```

```{r}
eval(x)
```

```{r}
evalq(x)
```
```{r}
eval(quote(x))
```

## 2.13.1 Lexical scope. Likelihoods

```{r}
Rmlfun = function(x) {
  sumx = sum(x)
    n = length(x)
    function(mu) n * log(mu) - mu * sumx
}

efun = Rmlfun(1:10)
efun(3)
```

```{r}
efun2 = Rmlfun(20:30)
efun2(3)
```

```{r}
efun = Rmlfun(1:10)
efun(3)
```

