---
title: "S4_05: the seed trap example"
output: html_notebook
---

Source: book

First, we tell R about the class itself, using the setClass function, which
takes the proposed class name and the proposed structure of the objects of
the class as arguments.

```{r}
setClass("trapTransect", representation(distances = "numeric", 
                                        seed.counts = "numeric",
                                        trap.area = "numeric"))
```

Writing an object constructor is a little more involved than for S3 classes.
The constructor function is called `new`, but if we wish to do any processing
of the arguments, including validity checks, then we need to add a speciﬁc
initialize function, which will be called by new.

```{r}
setMethod("initialize", "trapTransect", 
          function(.Object, 
                   distances = numeric(0), 
                   seed.counts = numeric(0), 
                   trap.area = numeric(0)) {
                if (length(distances) != length(seed.counts))
                    stop("Lengths of distances and counts differ.")
                if (length(trap.area) != 1)
                    stop("Ambiguous trap area.")
               .Object@distances <- distances
               .Object@seed.counts <- seed.counts
               .Object@trap.area <- trap.area
               .Object
               })
```

`new` creates an empty object and passes it to initialize, along with the
arguments that were provided to it. initialize then returns the updated
object, if the evaluations are successful.

```{r}
s1 <- new("trapTransect",
                         distances = 1:4,
                         seed.counts = c(4, 3, 2, 0),
                         trap.area = 0.0001)
```

```{r}
initialize(s1, 1, 3, 0.001)
```



Objects from S4 classes diﬀer from objects of S3 classes in a few important
ways. The elements that comprise the object, as deﬁned in the setClass
function, are called slots. The names of the slots can be found by

```{r}
slotNames(s1)
```

The values in the slots are accessed by either the slot function or the “@”
operator, which takes the place of the $ operator used previously.

```{r}
s1@distances
```

We now add two methods for the class: `show`, to print objects of the class when
just the object name is input, and `mean`, to compute and return the mean seed
distance from the object. In each case we use the `setMethod` function, which
requires the method name, the pattern of expected formal arguments (called
the signature), and the function itself.

```{r}
setMethod("mean", 
          signature(x = "trapTransect"), 
          function(x, ...) 
              weighted.mean(x@distances, w = x@seed.counts))
```

```{r}
setMethod("show", 
          signature(object = "trapTransect"), 
          function(object) 
              str(object))
```

We demonstrate the application of the new methods to the object.

```{r}
s1
```

```{r}
mean(s1)
```

We list the S4 methods for the trapTransect class by

```{r}
showMethods(classes = "trapTransect")
```

To display the code for a particular S4 method, we use

```{r}
 getMethod("mean", "trapTransect")
```

