---
title: "R Notebook"
output: html_notebook
---

```{r}
setClass("Passenger", 
         representation(name = "character", 
                        origin = "character", 
                        destination = "character"))

```

```{r}
setClass("FreqFlyer", 
         representation(ffnumber = "numeric"), 
         contains = "Passenger")
```


```{r}
getClass("FreqFlyer")
```

```{r}
subClassMetaName("Passenger")
```

```{r}
superClassNames("FreqFlyer")
```

## 3.2.3 Abstract data types

Consider the following simple example, in S4. Suppose that we have a
Rectangle class and that this class should respond to requests that ask for
the area of the rectangle. We ﬁrst deﬁne a class for rectangle that includes
a speciﬁc slot for the area. We next deﬁne a generic function for area, and
deﬁne a method for the Rectangle class.

```{r}
setClass("Rectangle", representation(h = "numeric", 
                                     w = "numeric", 
                                     area = "numeric"))

myr = new("Rectangle", h = 10, w = 20, area = 200)

setGeneric("area", function(shape) standardGeneric("area"))
```

```{r}
setMethod("area", signature(shape = "Rectangle"), 
          function(shape) shape@area)
```

```{r}
myr@area
```

```{r}
area(myr)
```

Any user can either access the area directly by accessing the slot with
myr@area or by calling the area generic function. Accessing the slot breaks
the data type abstraction; you are relying on the implementation. Using the
generic function makes use of the ADT. If the representation were to change
to that shown below, any code relying on the generic function will continue
to work and any code relying on slot access will fail. By using ADTs, it is
simpler to change the representation of data types as a project evolves.

```{r}
setClass("Rectangle", representation(h = "numeric", w = "numeric"))
setMethod("area", "Rectangle", function(shape) shape@h * shape@w)

```

```{r}
myr = new("Rectangle", h = 15, w = 25)
area(myr)
```

## 3.2.4 Self-describing data

One of the major uses of OOP within the Bioconductor Project is in the
construction of self-describing data classes. The most widely used is the
ExpressionSet class deﬁned in the Biobase package. Our goal is to deﬁne
a self-describing data object that can be used to carry out a reasonable anal-
ysis of the data. If all information is stored in a single object it is easier to
save it, to share it with others, or to use it as input to a function. This is
consistent with the notion that you would like to place all data relevant to
the experiment into a single ﬁle folder and to place it into a ﬁling cabinet so
that later you can ﬁnd all the information you need in one place.

The data might be stored in either a matrix or a data.frame. And while
informative row and column labels can be used, it is diﬃcult to encode all
relevant information about the variables in the labels. One solution is to
create a compound object that holds both the data and the metadata about
the variables, and possibly about the samples. Deﬁning a suitable class yields
self-describing data.

The major beneﬁts that we have found to programming with self-describing
data are that it is easy to return to a project after some months and re-do an
analysis. We have also found that it is relatively easy to hand oﬀ a project
from one analyst to another. But perhaps the greatest beneﬁt has come from
deﬁning specialized subsetting methods, that is, methods for [ that help to
construct an appropriate subset of the object, with all variables correctly
aligned.

```{r}
setClass("A", representation(s1 = "numeric"), 
         prototype = prototype(s1 = 0))

myA = new("A")
myA
```

```{r}
m2 = new("A", s1 = 10)
m2
```

We can create a second class B that contains A, so that B is a direct
subclass of A or, put another way, B inherits from class A. Any instance
of the class B will have all the slots in the A class and any additional ones
deﬁned speciﬁcally for B. Duplicate slot names are not allowed, so the slot
names for B must be distinct from those for A.

```{r}
setClass("B", contains = "A", 
         representation(s2 = "character"), 
         prototype = list(s2 = "hi"))

myB = new("B")
myB
```


```{r}
getSlots("A")
```

```{r}
slotNames("A")
```

```{r}
getClass("A")
```

```{r}
extends("B")
```

```{r}
extends("B", "A")

extends("A", "B")
```

Install first RBioinf with `BiocInstaller::biocLite("RBioinf")`

```{r}
RBioinf::superClassNames("B")
```

```{r}
RBioinf::subClassNames("A")
```

