---
title: "R Notebook"
output: html_notebook
---

The general idea is to be able to add elements to a container, and modify the
container in-place. In R this is possible using some assignment tricks into the
.GlobalEnv, but it can be frought with unintended consequences. Closures offer
us a perfect alternative to keep surprises to a minimum. First, we'll create
single environment that will act as the container and then add into that
environment a stack vector and the two methods, push and pop.

```{r}
s <- new.env()

  s$.Data <- vector()
  s$push <- function(x) .Data <<- c(.Data, x)
  s$pop  <- function() {
      tmp <- .Data[length(.Data)]
      .Data <<- .Data[-length(.Data)]
      return(tmp)
  }

ls(s)
ls(s, all = TRUE)             # shot .dot objects
# [1] ".Data" "pop"   "push" 
```

We are using the double arrow <<- assignment operator in the push function to
let assignment proceed up the internal stack frame until a variable is found to
bind to. This allows for non-local modifications to our .Data variable. The push
method appends new data to the stack and pop removes the last element of the
stack and returns it to the caller. We can use the $ operator to access the
internal methods of our environment.



```{r}
s$push(1)
# Error in s$push(1) : object '.Data' not found
```

Oops, something is wrong. It turns out that <<– can't find the .Data object
stored in the s object. We haven't matched the environment of the function to
the object's environment. R isn't starting its search for .Data in the correct
location; it needs more information. The functions environment and
as.environment work well here.

```{r}
environment(s$push) <- as.environment(s)
environment(s$pop)  <- as.environment(s)

s$push(1)   # works now
s$.Data
# [1] 1
```

```{r}
s$pop()
s$.Data
# [1] 1
# numeric(0)
```



```{r}
new_stack <- function() { 
        stack <- new.env()
        stack$.Data <- vector()
        stack$push <- function(x) .Data <<- c(.Data,x)
        stack$pop  <- function() {
          tmp <- .Data[length(.Data)]
          .Data <<- .Data[-length(.Data)]
          return(tmp)
        }
        environment(stack$push) <- as.environment(stack)
        environment(stack$pop) <- as.environment(stack)
        class(stack) <- "stack"
        stack
}

  new_betterstack <- function() {
        stack <- new_stack()
        stack_env <- as.environment(stack)
        stack$shift   <- function(x) .Data <<- c(x, .Data)
        stack$unshift <- function() {
          tmp <- .Data[1]
          .Data <<- .Data[-1]
          return(tmp)
        }
        environment(stack$shift)   <- stack_env
        environment(stack$unshift) <- stack_env
        class(stack) <- c("betterstack", "stack")
        stack
  }
  
push <- function(x, value, ...) UseMethod("push")
pop  <- function(x, ...) UseMethod("pop")  
push.stack <- function(x, value, ...) x$push(value)
pop.stack  <- function(x) x$pop()

shift <- function(x, value, ...) UseMethod("shift")
unshift  <- function(x, ...) UseMethod("unshift")  
shift.stack <- function(x, value, ...) x$shift(value)
unshift.stack  <- function(x) x$unshift()
```

```{r}
nb <- new_betterstack()
push(nb, 1:3)
```

```{r}
nb$.Data
```

```{r}
pop(nb)
```

```{r}
 unshift(nb) # from the front
```

```{r}
shift(nb, 3)
```

```{r}
push(nb, 1)
```

```{r}
nb$.Data
```

## one module that contains push, pop, shift and unshift

```{r}
# making a solid module from new_stack and new_betterstack
flexStack <- function() {
    stack <- new.env()
    stack$.Data <- vector()
    stack$push <- function(x) .Data <<- c(.Data, x)
    stack$pop  <- function() {
        tmp <- .Data[length(.Data)]
        .Data <<- .Data[-length(.Data)]
        return(tmp)
    }
    stack$shift <- function(x) .Data <<- c(x, .Data)
    stack$unshift <- function() {
        tmp    <- .Data[1]
        .Data <<- .Data[-1]
        return(tmp)
    }
    environment(stack$push)      <- as.environment(stack)
    environment(stack$pop)       <- as.environment(stack)
    environment(stack$shift)     <- as.environment(stack)
    environment(stack$unshift)   <- as.environment(stack)    
    
    class(stack) <- "stack"
    stack
}

push <- function(x, value, ...) UseMethod("push")
pop  <- function(x, ...) UseMethod("pop")  
push.stack <- function(x, value, ...) x$push(value)
pop.stack  <- function(x) x$pop()

shift <- function(x, value, ...) UseMethod("shift")
unshift  <- function(x, ...) UseMethod("unshift")  
shift.stack <- function(x, value, ...) x$shift(value)
unshift.stack  <- function(x) x$unshift()

```

```{r}
setClass("stackS4", slots = c(
    .data = "numeric",
    stack = "environment"
)
# , prototype = prototype(
#     .data = vector("numeric"),
#     stack = new.env()
# )
)

setGeneric("push", function(object, value, ...) standardGeneric("push"))
setGeneric("pop", function(object, ...) standardGeneric("pop"))
setGeneric("shift", function(object, value, ...) standardGeneric("shift"))
setGeneric("unshift", function(object, ...) standardGeneric("unshift"))

setMethod("initialize", "stackS4", function(.Object, ...) {
    .Object@.data <- vector("numeric")
    .Object@stack <- new.env()
    # environment(stack$push)      <- as.environment(stack)
    # environment(stack$pop)       <- as.environment(stack)
    # environment(stack$shift)     <- as.environment(stack)
    # environment(stack$unshift)   <- as.environment(stack)    
    
    # class(stack) <- .Object@stack
    # class(.Object@stack) <- .Object@stack
    # stack    
    .Object
})

setMethod("push", "stackS4", function(object, value, ...) {
    # environment(object)      <- as.environment(object@stack)
    object@.data <- c(object@.data, value)
    object
})

setMethod("pop", "stackS4", function(object, ...) {
        # environment(object)      <- as.environment(object@stack)
        tmp <- object@.data[length(object@.data)]
        object@.data <- object@.data[-length(object@.data)]
        return(tmp)
})

st <- new("stackS4")
st
# nb <- new_betterstack()
push(st, 1:3)
```

```{r}
pop(st, 3)
st@.data
```






```{r}
nb <- flexStack()
push(nb, 1:3)
```

```{r}
nb$.Data
# [1] 1 2 3
```

```{r}
pop(nb)        # from the back
nb$.Data
# [1] 3
# [1] 1 2
```

```{r}
unshift(nb) # from the front
# [1] 2
```

```{r}
shift(nb, 3)
```

```{r}
push(nb, 1)
```

```{r}
nb$.Data
```



```{r}
setClass('Stack', 
         representation(list='list', cursor='numeric'),  # type defs
         prototype(list=list(), cursor=NA_real_))        # default values

setGeneric('push', function(obj, ...) standardGeneric('push'))
setMethod('push', signature(obj='Stack'), 
    function(obj, x) {
        obj@list <- c(x, obj@list)
        obj
})

setGeneric('pop', function(obj, ...) standardGeneric('pop'))
setMethod('pop', signature(obj='Stack'),
    function(obj) {
        obj@cursor <- obj@list[[1]]
        obj@list <- obj@list[-1]
        obj
    }
)


```

