---
title: "R Notebook"
output: html_notebook
---

In this section we look at another example, and we examine some of the functions
associated with S4 classes. The example we define will be used to motivate the
use of methods associated with a class, and it will be used to demonstrate
inheritance later. The idea is that we want to create a program to simulate a
cellular automata model of a predator-prey system.

We do not develop the whole code here but concentrate on the data structures. In
particular we will create a base class for the agents. In the next section we
will create the basic methods for the class. In the inheritance section we will
discuss how to build on the class to create different predators and different
prey species. The basic structure of the class is shown in Figure 1.

The methods for this class are defined in the following section. Here we define
the class and its slots, and the code to define the class is given below:

```{r}
######################################################################
# Create the base Agent class
#
# This is used to represent the most basic agent in a simulation.
Agent <- setClass(
        # Set the name for the class
        "Agent",

        # Define the slots
        slots = c(
                location = "numeric",
                velocity   = "numeric",
                active   = "logical"
                ),

        # Set the default values for the slots. (optional)
        prototype = list(
                location = c(0.0,0.0),
                active   = TRUE,
                velocity = c(0.0,0.0)
                ),

        # Make a function that can test to see if the data is consistent.
        # This is not called if you have an initialize function defined!
        validity = function(object) {
                if(sum(object@velocity^2)>100.0) {
                        return("The velocity level is out of bounds.")
                }
                return(TRUE)
        })
```

Now that the code to define the class is given we can create an object whose class is Agent.


```{r}
a <- Agent()
a
```

Before we define the methods for the class a number of additional commands are
explored. The first set of functions explored are the `is.object` and the `isS4`
commands. The `is.object` command determines whether or not a variable refers to
an object. The `isS4` command determines whether or not the variable is an S4
object. The reason both are required is that the isS4 command alone cannot
determine if a variable is an S3 object. You need to determine if the variable
is an object and then decide if it is S4 or not.

```{r}
is.object(a)
isS4(a)
```

The next set of commands are used to get information about the data elements, or
slots, within an object. The first is the `slotNames` command. This command can
take either an object or the name of a class. It returns the names of the slots
associated with the class as strings.

```{r}
slotNames(a)
slotNames("Agent")
```

The `getSlots` command is similar to the `slotNames` command. It takes the name of a
class as a string. It returns a vector whose entries are the types associated
with the slots, and the names of the entries are the names of the slots.

```{r}
getSlots("Agent")
s <- getSlots("Agent")
s[1]
s[[1]]
names(s)
```

The next command examined is the `getClass` command. It has two forms. If you give
it a variable that is an S4 class it returns a list of slots for the class
associated with the variable. If you give it a character string with the name of
a class it gives the slots and their data types.

```{r}
getClass(a)
getClass("Agent")
```

The final command examined is the `slot command`. It can be used to get or set the
value of a slot in an object. It can be used in place of the “@” operator.

```{r}
slot(a,"location")              # get the value of location
slot(a, "location") <- c(1, 5)  # set the value in location
a
```

## 2.3. Creating Methods
We now build on the Agent class defined above. Once the class and its data
elements are defined we can define the methods associated with the class. The
basic idea is that if the name of a function has not been defined, the name must
first be reserved using the `setGeneric` function. The `setMethod` can then be used
to define which function is called based on the class names of the objects sent
to it.

We define the methods associated with the Agent method given in the previous
section. Note that the `validity` function for an object is only called when it is
first created and when an explicit call to the `validObject` function is made. We
make use of the `validObject` command in the methods below that are used to change
the value of a data element within an object.

```{r}
# create a method to assign the value of the location
setGeneric(name="setLocation", def = function(theObject,position) {
                               standardGeneric("setLocation")
                       })

setMethod(f="setLocation", signature="Agent", definition=function(theObject,position) {
                              theObject@location <- position
                              validObject(theObject)
                              return(theObject)
                      })

# create a method to get the value of the location
setGeneric(name="getLocation", def=function(theObject) {
                               standardGeneric("getLocation")
                       })

setMethod(f="getLocation", signature="Agent", definition=function(theObject) {
                              return(theObject@location)
                      })


# create a method to assign the value of active
setGeneric(name="setActive", def=function(theObject,active) {
                               standardGeneric("setActive")
                       })

setMethod(f="setActive", signature="Agent", definition=function(theObject,active) {
                              theObject@active <- active
                              validObject(theObject)
                              return(theObject)
                      })

# create a method to get the value of active
setGeneric(name="getActive", def=function(theObject) {
                               standardGeneric("getActive")
                       })

setMethod(f="getActive", signature="Agent", definition=function(theObject) {
                              return(theObject@active)
                      })


# create a method to assign the value of velocity
setGeneric(name="setVelocity", def=function(theObject,velocity) {
                               standardGeneric("setVelocity")
                       })

setMethod(f="setVelocity", signature="Agent", definition=function(theObject,velocity) {
                              theObject@velocity <- velocity
                              validObject(theObject)
                              return(theObject)
                      })

# create a method to get the value of the velocity
setGeneric(name="getVelocity", def=function(theObject) {
                               standardGeneric("getVelocity")
                       })

setMethod(f="getVelocity", signature="Agent", definition=function(theObject) {
                              return(theObject@velocity)
                      })
```

With these definitions the data elements are encapsulated and can be accessed
and set using the methods given above. It is generally good practice in object
oriented programming to keep your data private and not show them to everybody
willy nilly.

```{r}
a <- Agent()
getVelocity(a)

a <-  setVelocity(a, c(1, 2))
getVelocity(a)
```

The last topic examined is the idea of overloading functions. In the examples
above the signature is set to a single element. The signature is a vector of
characters and specifies the data types of the argument list for the method to
be defined. Here we create two new methods. The name of the method is
`resetActivity`, and there are two versions.

The first version accepts two arguments whose types are `Agent` and logical. This
version of the method will set the activity slot to a given value. The second
version accepts two arguments whose types are Agent and numeric. This version
will set the activity to TRUE and then set the energy level to the value passed
to it. Note that the names of the variables in the argument list must be exactly
the same.

```{r}
# create a method to reset the velocity and the activity
setGeneric(name = "resetActivity", def = function(theObject,value) {
                               standardGeneric("resetActivity")
                       })

setMethod(f = "resetActivity", 
          signature = c("Agent","logical"), 
          definition = function(theObject,value) { 
              theObject <- setActive(theObject,value) 
              theObject <- setVelocity(theObject,c(0.0,0.0)) 
              return(theObject)
              })

setMethod(f="resetActivity", 
          signature = c("Agent","numeric"), 
          definition = function(theObject,value) { 
              theObject <- setActive(theObject,TRUE)
              theObject <- setVelocity(theObject,value) 
              return(theObject)
              })
```

This definition of the function yields two options for the resetActivity
function. The decision to determine which function to call depends on two
arguments and their type. For example, if the first argument is from the Agent
class and the second is a value of TRUE or FALSE, then the first version of the
function is called. Otherwise, if the second argument is a number the second
version of the function is called.

```{r}
a <- Agent()
a

a <- resetActivity(a,FALSE)
getActive(a)

a <- resetActivity(a,c(1,3))
getVelocity(a)
```

## 2.4. Inheritance
A class’ inheritance hiearchy can be specified when the class is defined using
the contains option. The contains option is a vector that lists the classes the
new class inherits from. In the following example we build on the Agent class
defined in the previous section. The idea is that we need agents that represent
a predator and two prey. We will focus on two predators for this example.

The hierarchy for the classes is shown in Figure 2.. In this example we have one
Prey class that is derived from the Agent class. There are two predator classes,
Bobcat and Lynx. The Bobcat class is derived from the Agent class, and the Lynx
class is derived from the Bobcat class. We will keep this very simple, and the
only methods associated with the new classes is a move method. For our purposes
it will only print out a message and set the values of the position and velocity
to demonstrate the order of execution of the methods associated with the
classes.

The first step is to create the three new classes.

```{r}
######################################################################
# Create the Prey class
#
# This is used to represent a prey animal
Prey <- setClass(
        # Set the name for the class
        "Prey",

        # Define the slots - in this case it is empty...
        slots = character(0),

        # Set the default values for the slots. (optional)
        prototype = list(),

        # Make a function that can test to see if the data is consistent.
        # This is not called if you have an initialize function defined!
        validity = function(object) {
                if(sum(object@velocity^2)>70.0) {
                        return("The velocity level is out of bounds.")
                }
                return(TRUE)
        },

        # Set the inheritance for this class
        contains = "Agent"
        )



######################################################################
# Create the Bobcat class
#
# This is used to represent a smaller predator
Bobcat <- setClass(
        # Set the name for the class
        "Bobcat",

        # Define the slots - in this case it is empty...
        slots = character(0),

        # Set the default values for the slots. (optional)
        prototype=list(),

        # Make a function that can test to see if the data is consistent.
        # This is not called if you have an initialize function defined!
        validity=function(object) {
                if(sum(object@velocity^2)>85.0) {
                        return("The velocity level is out of bounds.")
                }
                return(TRUE)
        },

        # Set the inheritance for this class
        contains = "Agent"
        )

######################################################################
# Create the Lynx class
#
# This is used to represent a larger predator
Lynx <- setClass(
        # Set the name for the class
        "Lynx",

        # Define the slots - in this case it is empty...
        slots = character(0),

        # Set the default values for the slots. (optional)
        prototype = list(),

        # Make a function that can test to see if the data is consistent.
        # This is not called if you have an initialize function defined!
        validity = function(object) {
                if(sum(object@velocity^2)>95.0) {
                        return("The velocity level is out of bounds.")
                }
                return(TRUE)
        },

        # Set the inheritance for this class
        contains = "Bobcat"
        )
```

The inheritance is specified using the contains option in the setClass command.
Note that this can be a vector allowing for multiple inheritance. We choose not
to use that to keep things simpler. If you are feeling like you need more
self-loathing in your life you should try it out and experiment.

Next we define a method, move, for the new classes. We will include methods for
the Agent, Prey, Bobcat, and Lynx classes. The methods do not really do anything
but are used to demonstrate the idea of how methods are executed.

```{r}
# create a method to move the agent.
setGeneric(name="move", def = function(theObject) {
                               standardGeneric("move")
                       })

setMethod(f="move",
                      signature="Agent",
                      definition=function(theObject)
                      {
                              print("Move this Agent dude")
                              theObject <- setVelocity(theObject,c(1,2))
                              validObject(theObject)
                              return(theObject)
                      }
                      )

setMethod(f="move",
                      signature="Prey",
                      definition=function(theObject)
                      {
                              print("Check this Prey before moving this dude")
                              theObject <- callNextMethod(theObject)
                              print("Move this Prey dude")
                              validObject(theObject)
                              return(theObject)
                      }
                      )

setMethod(f="move",
                      signature="Bobcat",
                      definition=function(theObject)
                      {
                              print("Check this Bobcat before moving this dude")
                              theObject <- setLocation(theObject,c(2,3))
                              theObject <- callNextMethod(theObject)
                              print("Move this Bobcat dude")
                              validObject(theObject)
                              return(theObject)
                      }
                      )

setMethod(f="move",
                      signature="Lynx",
                      definition=function(theObject)
                      {
                              print("Check this Lynx before moving this dude")
                              theObject <- setActive(theObject,FALSE)
                                theObject <- callNextMethod(theObject)
                              print("Move this Lynx dude")
                              validObject(theObject)
                              return(theObject)
                      }
                      )
```

There are a number of things to note. First each method calls the `callNextMethod`
command. This command will execute the next version of the same method for the
previous class in the hierarchy. Note that I have included the arguments (in the
same order) as those called by the original function. Also note that the
function returns a copy of the object and is used to update the object passed to
the original function.

Another thing to note is that the methods associated with the Lync, Bobcat, and
Agent classes arbitrarily change the values of the position, velocity, and
activity for the given object. This is done to demonstrate the changes that take
place and reinforce the necessity for using the callNextMethod function the way
it is used here.

Finally, it should be noted that the `validObject` command is called in every
method. You should try adding a print statement in the validity function. You
might find that the order is a bit odd. You should experiment with this and play
with it. There are times you do not get the expected results so be careful!

We now give a brief example to demonstrate the order that the functions are
called. In the example we create a Bobcat object and then call the move method.
We next create a Lynx object and do the same. We print out the slots for both
agents just to demonstrate the values that are changed.

```{r}
robert <- Bobcat()
robert

robert <- move(robert)
robert

lionel <- Lynx()
lionel

lionel <- move(lionel)
lionel
```

