--- title: "R Notebook" output: html_document: default html_notebook: default 
---

The following example shows how one can create quite complicated data structures
by using list objects as components inside a list.

```{r} spec.list <- vector(mode = "list", length = 3)

names(spec.list) <- c("O2", "H2O", "CO2")

spec.list$O2  <- list(name = "oxygen", mol.weight = 32) spec.list$H2O <-
list(name = "water", mol.weight = 18) spec.list$CO2 <- list(name = "carbon
dioxide", mol.weight = 44)

spec.list ``` The names() command initializes the names of the mother list. Each
daughter component of the mother list is in itself again a list object. Print
spec.list on the screen and try to understand its contents.

## Plotting a mathematical function

```{r} # 1st plot curve(expr = sin(3*pi*x))

# 2nd set of plots curve(expr = sin(3*pi*x), from = 0, to = 2, col = "blue", 
xlab = "x", ylab = "f(x)", main = "curve")

curve(expr = cos(3*pi*x), add = TRUE, col = "red", lty = 2) abline(h = 0, lty =
2)

# add legend legend("bottomleft", c("sin", "cos"), text.col = c("blue", "red"),
lty = 1:2) ```

## Root of functions

```{r} curve(expr = cos(x)-2*x, from = -2, to = 2) abline(h = 0, lty = 2) ```

## 9.2. The bisection method The bisection method is a basic root-ﬁnding method.
It is simple and robust, but also rather slow. In real life problems, other
methods will be preferred that are numerically more eﬃcient. However, the
bisection method nicely illustrates how a root algotithm works, and for this
reason, we discuss it here. We start from an interval [a, b] for which we know
that it contains a root. The bisection method then divides this interval in two
and selects the subinterval in which a root must lie. This procedure is repeated
until the interval is suﬃciently small (i.e. smaller than some preset
tolerance). The bisection method can be implented by following function:

```{r} bisection <- function(f, interval, tol=.Machine$double.eps^0.25,
maxiter=1000) { a <- interval[1] b <- interval[2] if (a >= b) stop("startpoint a
must be smaller than endpoint b") if (f(a)*f(b) >= 0) stop("f(a) and f(b) must
have a different sign") i <- 1 while (i <= maxiter) #limit iterations to prevent
infinite loop { mid <- (a + b)/2 # new midpoint estim.prec <- (b - a)/2 #
estimated precision if (f(mid) == 0 | estim.prec < tol) # root found { if
(f(mid) == 0) estim.prec <- 0 result <-
list(root=mid,f.root=f(mid),iter=i,estim.prec=estim.prec) return(result) } i <-
i + 1 # step counter increases if (sign(f(mid)) == sign(f(a))) a <- mid else b
<- mid } stop(paste("Bisection method failed: max number of
steps",maxiter,"exceeded")) } ```

The bisection method requires four diﬀerent types of input: * f: the function
for which the root needs to be found * interval: the interval speciﬁed by the
start point a and the end point b. The start point must be smaller than the end
point, and the function f needs to have an opposite sign at these two points. *
tol: the requested tolerance on the root. If no value is speciﬁed, we use a
default value based on the machine precision .Machine$double.eps * maxiter: the
maximum number of iterations. This criterion is speciﬁed to prevent that the
algorithm gets stuck in an inﬁnite loop.


The bisection method generates four diﬀerent types of output: * root: the
estimated value of the root * f.root: the function value evaluated at the root *
iter: the number of iterations used * estim.prec: an estimate of the precision
for the root ```{r} test.f <- function(x) return(cos(x)-2*x) (result <-
bisection(f = test.f, interval = c(-2, 2))) ```


## 11. Diﬀerential equations

In the natural sciences, the challenge is often to describe or predict how a 
given environmental variable will evolve over time. Depending on the problem at 
hand, this variable can have many diﬀerent meanings. For example, one could be 
interested in future evolution of the temperature of the atmosphere (as in the 
IPCC models of global warming), the seasonal ﬂuctuation in the oxygen 
concentration in a lake (as described in biogeochemical ecosystem models), or 
the year-to-year ﬂuctuations in the density of lions in the Serengeti National 
park (as in the population models used by conservation biologists). The 
mathematical models that describe such changes in environmental variable are 
called dynamic models, because they have the time t as the independent variable.
The models themselves are based on diﬀerential equations, which are typically of
the form:

$$\frac {dC}{dt} = f(a, b(t), C = f(C, t)$$ In this, $t$ represents the time and
$C$ is the `state variable`. Furthermore, $a$ is a parameter that remains
constant in time, while $b(t)$ is a function that varies with time (a so-called
**forcing function**). This equation is termed a diﬀerential equation because it
contains the derivative $dC/dt$, which represents the so-called `rate of change`
of the `state variable` $C$ . The speciﬁc form of the function f depends on the
physical, chemical and biological processes that inﬂuence $C$.

The determination of how the function $f$ should look like, is the subject of
the process of model development. In essence, one or more suitable diﬀerential 
equations need to be derived from the general balance equations of mass, 
momentum and/or energy, which form the theoretical pillars of any model in the 
natural sciences. How this is exactly done will not be considered here, but is 
treated in detail in the course on Environmental Modelling.






## 11.2. Types of diﬀerential equations * Autonomous versus non-autonomous
diﬀerential equations * Linear versus non-linear diﬀerential equations *
Ordinary versus partial diﬀerential equations

### Ordinary versus partial diﬀerential equations When the function $f$ only
contains the time $t$ and the state variable $C$ , the resulting equation is
called an **ordinary diﬀerential equation (ODE)**. When the function $f$ also
contains one or more diﬀerentials of the state variable $C$ with respect to a
spatial coordinate, the resulting equation is called a **partial diﬀerential
equation (PDE)**.

The use of `ODE`’s versus `PDE`’s to describe a natural system depends on the
level of detail one wants to include in the model. For example, the
concentration in a lake can be described with varying spatial detail. If the
lake is strongly mixed, concentrations will not exhibit spatial gradients, and
so the lake can be considered as one homogeneously mixed volume. In this case,
the lake can be described by a single concentration $C(t)$, which only depends
on time. However, when the lake is not fully mixed, more complex descriptions of
the concentration in the lake are needed. Following cases are possible:

* 0D-model: $C(t)$ only dependent on time (ODE case), 

* 1D-model: $C(t, z)$ depends on time and the spatial coordinate (e.g. z = depth in a lake),

* 2D-model: $C(t, x, z)$ depends on time and two spatial coordinates (e.g. z =
depth, x = distance along the axis of a river) 

* 3D-model: $C(t, x, y, z)$ depends on time and three spatial coordinates (e.g.
full three- dimensional model of a lake).

$$\frac {dC}{dt} = D \cdot (C_{in} - C) - k \cdot C$$
$$\frac {dC}{dt} = K_z \cdot \frac {\mathrm{d^2}C}{\mathrm{d}z^2} - k \cdot C$$
$$\frac {dC}{dt} = K_x \cdot \frac {\mathrm{d^2}C}{\mathrm{d}x^2} +
K_y \cdot \frac {\mathrm{d^2}C}{\mathrm{d}y^2} +
K_z \cdot \frac {\mathrm{d^2}C}{\mathrm{d}z^2} -
k \cdot C$$

The ﬁrst equation is an ODE or ordinary diﬀerential equation and describes how the average concentration in the lake will change through time (0D-model). 

The second equation comprises a PDE or partial diﬀerential equation, and describes how the concentration will change with time as well as with the depth of the lake (1D-model). 

The third equation is also a PDE, but now describes the concentration varies in all three directions (full 3D-model).

Note that for the solution of a PDE problem, one needs a set of boundary conditions in
addition to the initial conditions. Here, we only focus on the solution of ODE problems.
Accordingly, the implementation of boundary conditions and the solution of PDEs is for a
more advanced course. Intrinsically however, it is similar to the solution of ODEs.

## Example
Let’s consider again pour lake model,
$$\frac {dC}{dt} = D \cdot (C_{in} - C) - k \cdot C \\
C(t = t_0) = C_0$$

This is a linear ODE with constant parameters, and for this speciﬁc type of ODEs, one can prove that an analytical solution must always exist.

To see how the concentration will evolve over time, one simply needs to plot the analytical solution in a graph.

```{r}
# Parameters
D<-2; C_in<-1; k<-1; C_0<-2

# Time sequence vector
time.seq <- seq(from = 0, to = 0.5, by = 0.1)

# Analytical solution as a vector expression
C.star <- D/(D+k)*C_in
C.an <- C.star + (C_0 - C.star )*exp(-(D+k)*time.seq)

# Plot solution
plot(time.seq, C.an, type="l", ylim=c(0,2), lwd =2, cex = 2,
xlab="time", ylab="state variable", main = "Analytical solution")
```

