

https://www.r-bloggers.com/using-closures-as-objects-in-r/


## Immutability
One thing that might surprise some readers (even though familiar with R) is we said almost all R objects are immutable. At first glance this doesn’t seem to be the case consider the following:

```{r}
a <- list()
print(a)

a$b <- 1
print(a)
```

The list “a” sure seemed to change. In fact it did not, this is an illusion foisted on you by R using some clever variable rebinding. Let’s look at that code more closely:

```{r}
library('pryr')
## a <- list()
print(address(a))
## [1] "0x1059c5dc0"
a$b <- 1
print(address(a))
## [1] "0x105230668"
```

R simulated a mutation or change on the object “a” by re-binding a new value
(the list with the extra argument) to the symbol “a” in the environment we were
executing in. We see this by the address change, the name “a” is no longer
referring to the same value. “Environment” is a computer science term meaning a
structure that binds variable names to values. R is very unusual in that most R
values are immutable and R environments are mutable (what value a variable
refers to get changed out from under you). At first glance R appears to be
adding an item to our list “a”, but in fact what is doing is changing the
variable name “a” to refer to an entirely new list that has one more element.


## Closures: “poor man’s objects”
There are some cases where you do want shared changing state. Garrett uses a
nice example of drawing cards, we will use a simple example of assigning
sequential IDs. Consider the following code:

```{r}
idSource <- function() {
  nextId <- 1
  list(nextID=function() { 
    r <- nextId
    nextId <<- nextId + 1
    r
  })
}

source <- idSource()
source$nextID()
## [1] 1
source$nextID()
## [1] 2
```

The idea is the following: in R a fresh environment (that is the structure
binding variable names to values) is created during function evaluation. Any
function created while evaluating our outer function has access to all variables
in this environment (this environment is what is called a closure). So any names
that appear free in the inner function (that is variable names that don’t have a
definition in the inner function) end up referring to variable in the this new
environment (or one of its parents if there is no name match). 

Since
environments are mutable re-binding values in this secret environment gives us
mutable slots. The first gotcha is the need to use <<- or assign() to effect
changes in the secret environment.

This behaves a lot more like what Java or Python programmer would expect from an
object and is fully idiomatic R. So if you want object-like behavior this is a
tempting way to get it.



```{r}
# trying to observe address
f <- function() {
  a <- list(b = 1)
  print(pryr::address(a))
  a$b <- 2
  print(pryr::address(a))
  a$b <- 3
  print(pryr::address(a))
}
f()

# timing
f1c <- function(n,verbose,shadow) {
  v <- 1:n
  vLast <- c()
  if(shadow) {
    vLast <- v
  }
  if(verbose) {
    print(pryr::address(v))
  }
  for(i in 1:n) {
    v[i] <- v[i]^2
    if(shadow) {
       vLast <- v
    }
    if(verbose) {
       print(pryr::address(v))
    }
  }
  c()
}

f1c(5,TRUE,FALSE)

print(system.time(f1c(30000,FALSE,FALSE)))
print(system.time(f1c(30000,FALSE,TRUE)))
```


```{r}
library('biglm')

datn <- function(n) {
   d <- data.frame(x=runif(n))
   d$y <- d$x>=runif(nrow(d))
   d
}

d <- datn(100000)
formula <- 'y~x'

fitter <- function(formula,d) {
  model <- bigglm(as.formula(formula),d,family=binomial(link='logit'))
  list(predict=function(newd) {predict(model,newdata=newd,type='response')[,1]})
}

model <- fitter(formula,d)
print(head(model$predict(d)))

sizeTest1 <- function(d) {
  model <- fitter(formula,d)
  length(serialize(model, NULL))
}

for(n in c(1000,10000,100000)) {
  print(sizeTest1(datn(n)))
}


```

```{r}
library('biglm')

datn <- function(n) {
   d <- data.frame(x=runif(n))
   d$y <- d$x>=runif(nrow(d))
   d
}

d <- datn(100000)
formula <- 'y~x'

sizeTest1 <- function(d) {
  model <- fitter(formula,d)
  length(serialize(model, NULL))
}


#' build a new funcion with a smaller environment
#' @param f input function
#' @param varaibles names we are allowing to be captured in the closere
#' @return new function with closure restricted to varaibles
#' @export
restrictEnvironment <- function(f, varList) {
  oldEnv <- environment(f)
  newEnv <- new.env(parent=parent.env(oldEnv))
  for(v in varList) {
    assign(v,get(v,envir=oldEnv),envir=newEnv)
  }
  environment(f) <- newEnv
  f
}

fitter <- function(formula,d) {
  model <- bigglm(as.formula(formula),d,family=binomial(link='logit'))
  
  model$family$variance <- c()
  model$family$dev.resids <- c()
  model$family$aic <- c()
  model$family$mu.eta <- c()
  model$family$initialize <- c()
  model$family$validmu <- c()
  model$family$valideta <- c()
  model$family$simulate <- c()
  
  environment(model$terms) <- new.env(parent=globalenv())
  list(predict = restrictEnvironment(function(newd) {
           predict(model, newdata = newd, type='response')[,1]
          }, 'model'))
}



model <- fitter(formula, d)
print(head(model$predict(d)))

for(n in c(1000,10000,100000)) {
  print(sizeTest1(datn(n)))
}
```



```{r}
library('biglm')
datn <- function(n) {
   d <- data.frame(x=runif(n))
   d$y <- d$x>=runif(nrow(d))
   d
}

d <- datn(100000)
formula <- 'y~x'

model <- bigglm(as.formula(formula),d,family=binomial(link='logit'))

```

